### 2019-02-28 ms.Rmd
Start conversion to drake plan

---
title: Growth Dynamics of the Independent Gametophytes of *Pleurorosiopsis makinoi* (Polypodiaceae)
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)

# Load packages
library(conflicted)
library(tidyverse)
library(drake)
library(here)
library(cowplot)
library(ggplot2)
library(scales)
library(here)
library(patchwork)
library(lubridate)
library(gt)
library(ggridges)

# Resolve conflicts
col_factor <- conflict_prefer("col_factor", "readr")
discard    <- conflict_prefer("discard", "purrr")
expand     <- conflict_prefer("expand", "tidyr")
filter     <- conflict_prefer("filter", "dplyr")
gather     <- conflict_prefer("gather", "tidyr")
ggsave     <- conflict_prefer("ggsave", "ggplot2")
here       <- conflict_prefer("here", "here")
intersect  <- conflict_prefer("intersect", "dplyr")
lag        <- conflict_prefer("lag", "dplyr")
Position   <- conflict_prefer("Position", "ggplot2")
setdiff    <- conflict_prefer("setdiff", "dplyr")
union      <- conflict_prefer("union", "dplyr")

source(here("code/functions.R"))

#' Set order of microclimate variables for printing.
order_vars <- c("par_total", "rh_min", "temp_mean")

#' Select microclimate variables of interest for analysis.
selected_vars <- c("rh_min", "par_total", "temp_mean")

# Load objects from drake workflow into memory
loadd(combined_morph)
loadd(daily_microclimate)
loadd(t_test_results)
loadd(paired_microclimate)
loadd(nested_paired_microclimate)
```

```{r fig-3, fig.width=8, fig.height=7}
#' Set x-limits manually since the automatically set limits will vary
#' when we subset the data.
start_date <- combined_morph %>%
  pull(date) %>%
  min

end_date <- combined_morph %>%
  pull(date) %>%
  max

#' Make tibble of years for shading. This dataframe will be passed to `geom_rect()` when
#' plotting.
shading_dates <- 
  tibble(date = seq.Date(from = start_date, to = end_date, by = "day")) %>%
  mutate(year = year(date)) %>%
  # shade by alternate years
  filter(year %in% c(2009, 2011, 2013)) %>%
  group_by(year) %>%
  summarize(
    year_start = min(date),
    year_end = max(date)
  )

#' Make list of subplots. Build each one at a time, then add common
#' formatting. 
subplots <- list()

subplots[["cover"]] <- 
  combined_morph %>%
  ungroup %>%
  select(starts_with("q"), date) %>%
  filter(complete.cases(.)) %>%
  gather(plot, area, -date) %>%
  ggplot(aes(x = date, y = area, linetype = plot)) %>%
  shade_years(shading_dates) +
  geom_line() +
  scale_linetype_manual(
    name="Quadrat",
    breaks=c("q_1", "q_2", "q_3", "q_4"),
    labels=c("1", "2", "3", "4"),
    values=c("solid", "dashed", "dotted", "dotdash")
  ) +
  labs(
    x = "",
    y = expression("Cover ("~cm^2~")"),
    subtitle = "a"
  ) +
  theme(legend.position = "bottom")

legend <- get_legend(subplots[["cover"]])

subplots[["cover"]] <- 
  subplots[["cover"]] +
  theme(legend.position = "none")

subplots[["gemmae_count"]] <- 
  combined_morph %>%
  select(count_mean, count_sd, date) %>%
  drop_na %>%
  ggplot(aes(x = date, y = count_mean)) %>%
  shade_years(shading_dates) +
  geom_errorbar(
    aes(ymin=count_mean-count_sd, 
        ymax=count_mean+count_sd), 
    width=.1,
    color = "dark grey") +
  geom_line() +
  geom_point(color = "black") +
  labs(
    x = "",
    y = expression("Gemmae count"),
    subtitle = "b"
  )

subplots[["gemmae_length"]] <- 
  combined_morph %>%
  select(length_mean, length_sd, date) %>%
  drop_na %>%
  ggplot(aes(x = date, y = length_mean)) %>%
  shade_years(shading_dates) +
  geom_errorbar(
    aes(ymin=length_mean-length_sd, 
        ymax=length_mean+length_sd), 
    width=.1,
    color = "dark grey") +
  geom_line() +
  geom_point(color = "black") +
  labs(
    x = "",
    y = expression(paste("Gemmae len. (", mu, "m)")),
    subtitle = "c"
  )

subplots <- subplots %>%
  map(~ . + 
        theme(
          axis.text.x = element_text(
            angle = 30, 
            hjust = 1, 
            vjust = 0.5, 
            margin=margin(-10,0,0,0)
          ),
          plot.margin = margin(0,10,0,10)
        ) +
        scale_x_date(
          date_labels = "%b %Y",
          date_breaks = "6 months",
          limits = c(start_date, end_date)
        )
  )

subplots[1:2] <- subplots[1:2] %>%
  map(~ . + theme(axis.text.x = element_blank()) 
  )

subplots[2:3] <- subplots[2:3] %>%
  map(~ . + theme(plot.margin = margin(-10,10,0,10)) 
  )

#' Combine subplots and write out
subplots[[1]] + subplots[[2]] + subplots[[3]] + 
  legend +
  plot_layout(ncol = 1, heights = c(1,1,1,0.2))

```

```{r fig-4, fig.width=8, fig.height=7}

#' Extract common start and end dates so x-axis are same across subplots.
start_date <- daily_microclimate %>% pull(date) %>% min
end_date <- daily_microclimate %>% pull(date) %>% max
all_days <- tibble( 
  date = seq(start_date, end_date, by = "day")
)

#' Make tibble of years for shading.
shading_dates <- 
  tibble(date = seq.Date(from = start_date, to = end_date, by = "day")) %>%
  mutate(year = year(date)) %>%
  # shade by odd years
  filter(year %in% c(2009, 2011, 2013)) %>%
  group_by(year) %>%
  summarize(
    year_start = min(date),
    year_end = max(date)
  )

#' PAR differs enough between sites that automatically set y-axis limits
#' would be different. Use the common min and max between them.
min_par <- daily_microclimate %>% pull(par_total) %>% min
max_par <- daily_microclimate %>% pull(par_total) %>% max

#' Insert NAs into microclimate for each site so that missing days
#' don't get connected by lines.
daily_microclimate_okutama <- filter(daily_microclimate, site == "okutama") %>%
  right_join(all_days)

daily_microclimate_takao <- filter(daily_microclimate, site == "uratakao") %>%
  right_join(all_days)

daily_microclimate_with_na <- bind_rows(
  daily_microclimate_okutama, daily_microclimate_takao)

#' Make list of subplots.
subplots <- list()

subplots[[1]] <- 
  daily_microclimate_okutama %>%
  ggplot(aes(x = date, y = par_total)) %>%
  shade_years(shading_dates) +
  geom_line() +
  scale_y_continuous(limits = c(min_par, max_par)) +
  labs(y = expression(paste("DLI (", mol~m^-2~day^-1, ")", sep = "") ),
       x = "",
       title = "Okutama",
       subtitle = "a")

subplots[[2]] <- 
  daily_microclimate_takao %>%
  ggplot(aes(x = date, y = par_total)) %>%
  shade_years(shading_dates) +
  geom_line() +
  scale_y_continuous(limits = c(min_par, max_par)) +
  labs(y = "",
       x = "",
       title = "Uratakao",
       subtitle = "b")

subplots[[3]] <- 
  daily_microclimate_okutama %>%
  ggplot(aes(x = date, y = rh_min)) %>%
  shade_years(shading_dates) +
  geom_line() +
  scale_y_continuous(limits = c(0, 100)) +
  labs(y = "Min. Rel. Humidity (%)",
       x = "",
       subtitle = "c")

subplots[[4]] <- 
  daily_microclimate_takao %>%
  ggplot(aes(x = date, y = rh_min)) %>%
  shade_years(shading_dates) +
  geom_line() +
  scale_y_continuous(limits = c(0, 100)) +
  labs(y = "",
       x = "",
       subtitle = "d")

subplots[[5]] <- 
  daily_microclimate_okutama %>%
  ggplot(aes(x = date, y = temp_mean)) %>%
  shade_years(shading_dates) +
  geom_line() +
  labs(y = "Mean Temp. (°C)",
       x = "",
       subtitle = "e")

subplots[[6]] <- 
  daily_microclimate_takao %>%
  ggplot(aes(x = date, y = temp_mean)) %>%
  shade_years(shading_dates) +
  geom_line() +
  labs(y = "",
       x = "",
       subtitle = "f")

#' Apply common formatting.
subplots <- subplots %>%
  map(~ . + 
        theme(
          axis.text.x = element_text(
            angle = 30, 
            hjust = 1, 
            vjust = 0.5, 
            margin=margin(-10,0,0,0)
          )
        ) +
        scale_x_date(
          date_labels = "%b %Y",
          date_breaks = "6 months",
          limits = c(start_date, end_date)
        )
  )

#' Remove x-axis lables from upper plots.
subplots[1:4] <- subplots[1:4] %>%
  map(~ . + theme(axis.text.x = element_blank()) 
  )

#' Close gaps on top margins for lower plots.
subplots[c(3,5)] <- subplots[c(3,5)] %>%
  map(~ . + theme(plot.margin = margin(-10,10,0,10)) 
  )

#' Close gaps on left side for RHS plots.
subplots[c(2,4,6)] <- subplots[c(2,4,6)] %>%
  map(~ . + theme(plot.margin = margin(-10,10,0,-40)) 
  )

#' Combine plots and write out.
subplots[[1]] + subplots[[2]] + subplots[[3]] + 
  subplots[[4]] + subplots[[5]] + subplots[[6]] + 
  plot_layout(ncol = 2)

```

```{r table-1}
#' Format Table 1 and output to RTF using `gt` package.
t_test_results %>%
  arrange(factor(var, levels = order_vars)) %>%
  mutate(season = stringr::str_to_title(season)) %>%
  mutate(var = case_when(
    var == "rh_min" ~    "Min. Rel. Humidity",
    var == "par_total" ~ "DLI",
    var == "temp_mean" ~ "Mean Temp."
  )) %>%
  group_by(season) %>%
  gt() %>%
  fmt_number(
    columns = vars(estimate, statistic, conf.low, conf.high, 
                   okutama_mean, okutama_sd, 
                   uratakao_mean, uratakao_sd)
  ) %>%
  cols_move(
    columns = vars(okutama_mean, okutama_sd),
    after = vars(conf.high)
  ) %>%
  tab_spanner(
    label = "Okutama",
    columns = vars(okutama_mean, okutama_sd)
  ) %>%
  tab_spanner(
    label = "Uratakao",
    columns = vars(uratakao_mean, uratakao_sd)
  ) %>%
  cols_label(
    var = "",
    statistic = "t",
    pval = "p",
    parameter = "df",
    conf.low = "Lower",
    conf.high = "Upper",
    okutama_mean = "Mean",
    okutama_sd = "SD",
    uratakao_mean = "Mean",
    uratakao_sd = "SD"
  ) %>%
  tab_style(
    style = cells_styles(
      bkgd_color = "white"),
    locations = cells_data(
      columns = everything()
    )
  ) %>%
  tab_header(
    title = "Table 1"
  )

```

```{r fig-5, fig.width=7, fig.height=8}
#' Extract p-values for plotting, and put into data frame
#' with xvals specifying season and yvals specifying height 
#' to plot asterisks.
p_vals <-
  t_test_results %>%
  mutate(pval = case_when(
    pval < 0.0001 ~ "***",
    pval < 0.001 ~ "**",
    pval < 0.05 ~ "*",
    TRUE ~ ""
  )) %>%
  select(season, var, pval) %>%
  spread(var, pval) %>%
  rename(
    par_total_pval = par_total,
    rh_min_pval = rh_min,
    temp_mean_pval = temp_mean
  )

# Simple function to calculate maximum y-position for plotting asterisks.
# Should be just above the mean + error bar height (1 sd).
maxy <- function(x) {mean(x, na.rm = TRUE) + sd(x, na.rm = TRUE)}

pval_plot_data <- 
  paired_microclimate %>%
  select(-date, -month) %>%
  group_by(season, site) %>%
  summarize_all(
    maxy
  ) %>%
  group_by(season) %>%
  summarize_all(
    max
  ) %>%
  select(-site) %>%
  left_join(p_vals)

#' Format data for plotting: need columns to be mean and sd, by site and season.
paired_means <- 
  paired_microclimate %>%
  select(-date) %>%
  group_by(site, season) %>%
  summarize_all(
    list(~mean(., na.rm = TRUE),
         ~sd(., na.rm = TRUE))
  ) %>%
  mutate(
    season = factor(season, levels = c("winter", "spring", "summer", "fall"))
  )

#' Make plot including t-test on three microclimatic variables.
#' 
#' The final plot includes multiple subplots. Each is a ridge plot showing 
#' the distribution of one of the variables between the two sites, with the
#' results of a paired t-test showing if they are different or not.
#' 
#' Need to manually set x ranges to keep x-axis scale the same across plots
x_ranges <- tibble(var = selected_vars) %>%
  mutate(
    min = map_dbl(var, ~ pull(paired_microclimate, .) %>% min ),
    max = map_dbl(var, ~ pull(paired_microclimate, .) %>% max )
  )

# light blue / light green colorblind-friendly from colorbrewer
# paired_colors <- c("#a6cee3", "#b2df8a")
# OR dark blue / dark green colorblind-friendly from colorbrewer
# paired_colors <- c("#1f78b4", "#33a02c")
# OR dark grey / light grey for bw
paired_colors <- c("grey55", "grey90")

subplots <- list()

#' Make ridge plots by mapping across nested datasets.
plot_results <-
  nested_paired_microclimate %>%
  right_join(x_ranges) %>%
  mutate(min = min - 0.05 * min,
         max = max + 0.05 * max,
         plot = pmap(list(var, data, x_min = min, x_max = max), 
                     plot_ridges, 
                     y_var = "site"),
         pval = map2(var, data, run_t_test),
         pval = case_when(
           pval < 0.0001 ~ "***",
           pval < 0.001 ~ "**",
           pval < 0.05 ~ "*",
           TRUE ~ ""
         ),
         plot = map2(
           plot, pval, 
           ~ .x + annotate("text", Inf, Inf, 
                           label = .y, hjust = 1, vjust = 1))

         # If needed, add annotation for season to make sure subplots
         # are positioned correctly

         # ,plot = map2(
         #   plot, season,
         #   ~ .x + annotate("text", -Inf, Inf,
         #                   label = .y, hjust = -1, vjust = 1)),
         # season = as_factor(season),
         # season = fct_relevel(season, "winter", "spring", "summer", "fall")

  ) %>% 
  arrange(season, var)

subplots <- plot_results$plot

#' Manually some subplots to remove redundant labels, etc.
subplots[1:9] <- map(subplots[1:9], 
                     ~ . + theme(axis.title.x = element_blank(),
                                 axis.text.x = element_blank() ))

subplots[1:12] <- map(subplots[1:12], 
                      ~ . + scale_fill_manual(
                        values = paired_colors,
                        name="Site",
                        breaks=c("okutama", "uratakao"),
                        labels=c("Okutama", "Uratakao"))
)

subplots[[10]] <- subplots[[10]] +
  labs(x = expression(paste("DLI (", mol~m^-2~day^-1, ")", sep = "")) )

subplots[[11]] <- subplots[[11]] +
  labs(x = expression("Min. Rel. Humidity (%)"))

subplots[[12]] <- subplots[[12]] +
  labs(x = expression("Mean Temp. (°C)"))

#' Add titles.
subplots <- map(subplots, ~ . + labs(title = "") )

subplots[[2]] <- subplots[[2]] +
  labs(title = "Winter\n")

subplots[[5]] <- subplots[[5]] +
  labs(title = "Spring\n")

subplots[[8]] <- subplots[[8]] +
  labs(title = "Summer\n")

subplots[[11]] <- subplots[[11]] +
  labs(title = "Fall\n")

#' Extract legend
legend <- paired_microclimate %>%
  ggplot(aes(x = temp_mean, y = site, fill = site)) +
  geom_density_ridges() +
  theme(legend.position = "bottom") +
  scale_fill_manual(
    values = paired_colors,
    name="Site",
    breaks=c("okutama", "uratakao"),
    labels=c("Okutama", "Uratakao"))

legend <- get_legend(legend)

#' Combine and write out.
subplots[[1]] +  subplots[[2]] +  subplots[[3]] + 
  subplots[[4]] +  subplots[[5]] +  subplots[[6]] + 
  subplots[[7]] +  subplots[[8]] +  subplots[[9]] + 
  subplots[[10]] + subplots[[11]] + subplots[[12]] +
  plot_spacer() + legend + plot_spacer() + plot_layout(ncol = 3, heights = c(1, 1, 1, 1, 0.2))
```

### 2019-02-28 code/drake_functions.R

@@ -1,0 +1,296 @@
+#' Clean microclimate data.
+#' 
+#' Add 12 hours if time is PM ("午後"). 
+#' Time units are in seconds when adding.
+#'
+#' @param raw_microclimate_data Tibble; Raw microclimate data read in from xlsx file.
+#' @return Tibble
+clean_microclimate <- function(raw_microclimate_data) {
+  raw_microclimate_data %>%
+  mutate(
+    pm = str_detect(date_time, "午後"),
+    date_time = mdy_hms(date_time, tz=Sys.timezone()),
+    date_time = case_when (
+      pm == TRUE ~ date_time + 60*60*12,
+      TRUE ~ date_time
+    ),
+    date = date(date_time)
+  ) %>%
+  select(-pm)
+  
+}
+
+#' Calculate daily microclimate values from clean data
+#' including multiple observerations per day.
+#' 
+#' For each day, calculates mean temperature (temp_mean), 
+#' minimum relative humidity (rh_mean), and total PAR (
+#' photosynthetically active radiation; par_total). Also adds
+#' columns for season and site name.
+#'
+#' @param site_1_data Tibble of microclimate data from one site.
+#' @param site_1_name String; name of first site.
+#' @param site_2_data Tibble of microclimate data from another site.
+#' @param site_2_name String; name of second site.
+#'
+#' @return Tibble
+calculate_daily_microclimate <- function (site_1_data, site_1_name, site_2_data, site_2_name) {
+  bind_rows(
+    mutate(okutama_microclimate, site = site_1_name),
+    mutate(takao_microclimate, site = site_2_name)
+  ) %>%
+  # only include days with data at both sites
+  filter(complete.cases(.)) %>% 
+  group_by(date, site) %>%
+  summarize(
+    temp_mean = mean(temp, na.rm = TRUE),
+    rh_min = min(rh, na.rm = TRUE),
+    par_total = sum(1800*par) / 1000000
+  ) %>%
+  ungroup %>%
+  # add season variable
+  mutate(
+    month = month(date),
+    season = case_when(
+      month >= 1 & month < 3 ~ "winter",
+      month >= 3 & month < 6 ~ "spring",
+      month >= 6 & month < 9 ~ "summer",
+      month >= 9 & month < 12 ~ "fall",
+      month == 12 ~ "winter"
+    )
+  )
+}
+
+
+#' Clean raw cover data and calculate total cover.
+#'
[…]

### 2019-02-28 code/packages.R

@@ -1,0 +1,33 @@
+# Load the packages this workflow depends on
+
+# Install ferncollectR (Fern collection data)
+# This is updated frequently, so try to install 
+# the newest package first. The pacakge is a private 
+# github repo, so need have GITHUB_PAT defined in 
+# .Renviron to access.
+
+library(conflicted)
+library(assertthat)
+library(assertr)
+library(drake)
+library(glue)
+library(here)
+library(jntools)
+library(lubridate)
+library(magrittr)
+library(readxl)
+library(tidyverse)
+library(broom)
+
+# Resolve conflicting functions
+collapse <- dplyr::collapse
+count <- dplyr::count
+expand <- tidyr::expand
+extract <- magrittr::extract
+filter <- dplyr::filter
+gather <- tidyr::gather
+intersect <- dplyr::intersect
+here <- here::here
+map <- purrr::map
+select <- dplyr::select
+setdiff <- dplyr::setdiff

### 2019-02-28 code/plan.R

@@ -1,0 +1,214 @@
+#' Workflow Plans
+#'
+#' ## Data sources
+#' 
+#' The raw data are in `.xlsx` files. Some of the data are pre-processed
+#' (mean and sd of count and length of gemmae, 30 min averages of PPFD). 
+#' 
+#' The files include some additional analsyes and other data that won't 
+#' be used here.
+
+data_cleaning_plan <- drake_plan (
+  
+  #' ### Microclimate data
+  #'
+  #' Microclimate variables include PPFD (photon flux density, in μmol of 
+  #' light per sq m per sec), rel. humidity (%), and temperature (°C) 
+  #' measured once every 30 min. 
+  #' 
+  #' PPFD is 30 min-averages of values taken every 4 minutes
+  #' (raw 4 minute values not included).
+  #' 
+  #' There are two sites, Okutama (site of the independent gametophyte
+  #' colony) and Uratakao (site of the sporophyte population). Additional
+  #' sites were also measured, but not included in this analysis
+  #' because of too much missing data due to mechanical failures.
+  #'
+  #' For Okutama, the raw data for temperature, rel. hum,
+  #' and light (PPFD) are in different columns
+  #' in the `xlsx` file, so read in each variable separately.
+  
+  okutama_temp_raw = read_excel(
+    file_in("data_raw/datalogger_raw.xlsm"),
+    sheet = 1,
+    range = "A2:B28683",
+    col_names = c("date_time", "temp")),
+  
+  okutama_rh_raw = read_excel(
+    file_in("data_raw/datalogger_raw.xlsm"),
+    sheet = 1,
+    range = "C2:D28683",
+    col_names = c("date_time", "rh")),
+  
+  okutama_par_raw = read_excel(
+    file_in("data_raw/datalogger_raw.xlsm"),
+    sheet = 1,
+    range = "F2:G28204",
+    col_names = c("date_time", "par")),
+  
+  #' Combine raw okutama microclimate variables into a single tibble.
+  okutama_microclimate_raw =
+    full_join(okutama_temp_raw, okutama_rh_raw) %>%
+    full_join(okutama_par_raw),
+  
+  #' Uratakao raw data have a single column for time of the three microclimate
+  #' variables, so these can be read in all at once.
+  takao_microclimate_raw = read_excel(
+    file_in("data_raw/datalogger_raw.xlsm"),
+    sheet = 2,
[…]

### 2019-02-28 make.R

@@ -1,0 +1,29 @@
+# make.R 
+
+# Master script for running T_angustifrons analyses.
+# This project uses drake to manage workflows.
+# For more information about drake, see
+# https://ropensci.github.io/drake/
+
+# Setup ----
+
+# Set working directory
+setwd(here::here())
+
+# Load packages
+source("code/packages.R")
+
+# Update drake settings
+pkgconfig::set_config("drake::strings_in_dots" = "literals")
+
+# Load functions and plans  ----
+
+# General functions
+source("code/functions.R")
+source("code/drake_functions.R")
+
+# Data processing
+source("code/plan.R")
+
+# Run analyses ----
+make(plan)

### 2019-03-01 code/drake_functions.R
Merge drake_functions into functions

@@ -1,296 +1,0 @@
-#' Clean microclimate data.
-#' 
-#' Add 12 hours if time is PM ("午後"). 
-#' Time units are in seconds when adding.
-#'
-#' @param raw_microclimate_data Tibble; Raw microclimate data read in from xlsx file.
-#' @return Tibble
-clean_microclimate <- function(raw_microclimate_data) {
-  raw_microclimate_data %>%
-  mutate(
-    pm = str_detect(date_time, "午後"),
-    date_time = mdy_hms(date_time, tz=Sys.timezone()),
-    date_time = case_when (
-      pm == TRUE ~ date_time + 60*60*12,
-      TRUE ~ date_time
-    ),
-    date = date(date_time)
-  ) %>%
-  select(-pm)
-  
-}
-
-#' Calculate daily microclimate values from clean data
-#' including multiple observerations per day.
-#' 
-#' For each day, calculates mean temperature (temp_mean), 
-#' minimum relative humidity (rh_mean), and total PAR (
-#' photosynthetically active radiation; par_total). Also adds
-#' columns for season and site name.
-#'
-#' @param site_1_data Tibble of microclimate data from one site.
-#' @param site_1_name String; name of first site.
-#' @param site_2_data Tibble of microclimate data from another site.
-#' @param site_2_name String; name of second site.
-#'
-#' @return Tibble
-calculate_daily_microclimate <- function (site_1_data, site_1_name, site_2_data, site_2_name) {
-  bind_rows(
-    mutate(okutama_microclimate, site = site_1_name),
-    mutate(takao_microclimate, site = site_2_name)
-  ) %>%
-  # only include days with data at both sites
-  filter(complete.cases(.)) %>% 
-  group_by(date, site) %>%
-  summarize(
-    temp_mean = mean(temp, na.rm = TRUE),
-    rh_min = min(rh, na.rm = TRUE),
-    par_total = sum(1800*par) / 1000000
-  ) %>%
-  ungroup %>%
-  # add season variable
-  mutate(
-    month = month(date),
-    season = case_when(
-      month >= 1 & month < 3 ~ "winter",
-      month >= 3 & month < 6 ~ "spring",
-      month >= 6 & month < 9 ~ "summer",
-      month >= 9 & month < 12 ~ "fall",
-      month == 12 ~ "winter"
-    )
-  )
-}
-
-
-#' Clean raw cover data and calculate total cover.
-#'
[…]

### 2019-03-01 code/functions.R

@@ -246,3 +246,438 @@ geom_rect(
   inherit.aes = FALSE
 )
 }
+
+#' Clean microclimate data.
+#' 
+#' Add 12 hours if time is PM ("午後"). 
+#' Time units are in seconds when adding.
+#'
+#' @param raw_microclimate_data Tibble; Raw microclimate data read in from xlsx file.
+#' @return Tibble
+clean_microclimate <- function(raw_microclimate_data) {
+  raw_microclimate_data %>%
+    mutate(
+      pm = str_detect(date_time, "午後"),
+      date_time = mdy_hms(date_time, tz=Sys.timezone()),
+      date_time = case_when (
+        pm == TRUE ~ date_time + 60*60*12,
+        TRUE ~ date_time
+      ),
+      date = date(date_time)
+    ) %>%
+    select(-pm)
+  
+}
+
+#' Calculate daily microclimate values from clean data
+#' including multiple observerations per day.
+#' 
+#' For each day, calculates mean temperature (temp_mean), 
+#' minimum relative humidity (rh_mean), and total PAR (
+#' photosynthetically active radiation; par_total). Also adds
+#' columns for season and site name.
+#'
+#' @param site_1_data Tibble of microclimate data from one site.
+#' @param site_1_name String; name of first site.
+#' @param site_2_data Tibble of microclimate data from another site.
+#' @param site_2_name String; name of second site.
+#'
+#' @return Tibble
+calculate_daily_microclimate <- function (site_1_data, site_1_name, site_2_data, site_2_name) {
+  bind_rows(
+    mutate(okutama_microclimate, site = site_1_name),
+    mutate(takao_microclimate, site = site_2_name)
+  ) %>%
+    # only include days with data at both sites
+    filter(complete.cases(.)) %>% 
+    group_by(date, site) %>%
+    summarize(
+      temp_mean = mean(temp, na.rm = TRUE),
+      rh_min = min(rh, na.rm = TRUE),
+      par_total = sum(1800*par) / 1000000
+    ) %>%
+    ungroup %>%
+    # add season variable
+    mutate(
+      month = month(date),
+      season = case_when(
+        month >= 1 & month < 3 ~ "winter",
+        month >= 3 & month < 6 ~ "spring",
+        month >= 6 & month < 9 ~ "summer",
+        month >= 9 & month < 12 ~ "fall",
+        month == 12 ~ "winter"
+      )
+    )
+}
+
+
+#' Clean raw cover data and calculate total cover.
+#'
[…]

### 2019-03-01 ms.Rmd
Finish running all analyses and output to ms as drake plan

@@
 
-# Load packages
-library(conflicted)
-library(tidyverse)
-library(drake)
-library(here)
-library(cowplot)
-library(ggplot2)
-library(scales)
-library(here)
-library(patchwork)
-library(lubridate)
-library(gt)
-library(ggridges)
-
-# Resolve conflicts
-col_factor <- conflict_prefer("col_factor", "readr")
-discard    <- conflict_prefer("discard", "purrr")
-expand     <- conflict_prefer("expand", "tidyr")
-filter     <- conflict_prefer("filter", "dplyr")
-gather     <- conflict_prefer("gather", "tidyr")
-ggsave     <- conflict_prefer("ggsave", "ggplot2")
-here       <- conflict_prefer("here", "here")
-intersect  <- conflict_prefer("intersect", "dplyr")
-lag        <- conflict_prefer("lag", "dplyr")
-Position   <- conflict_prefer("Position", "ggplot2")
-setdiff    <- conflict_prefer("setdiff", "dplyr")
-union      <- conflict_prefer("union", "dplyr")
-
-source(here("code/functions.R"))
-
-#' Set order of microclimate variables for printing.
+# Set order of microclimate variables for printing.
 order_vars <- c("par_total", "rh_min", "temp_mean")
 
-#' Select microclimate variables of interest for analysis.
-selected_vars <- c("rh_min", "par_total", "temp_mean")
-
 # Load objects from drake workflow into memory
+loadd(selected_vars)
 loadd(combined_morph)
@@
 loadd(nested_paired_microclimate)
+loadd(combined_monthly_morph)
+loadd(linear_model_results)
+loadd(gameto_size)
 ```
@@
 
-```{r fig-5, fig.width=7, fig.height=8}
-#' Extract p-values for plotting, and put into data frame
-#' with xvals specifying season and yvals specifying height 
-#' to plot asterisks.
-p_vals <-
-  t_test_results %>%
-  mutate(pval = case_when(
-    pval < 0.0001 ~ "***",
-    pval < 0.001 ~ "**",
-    pval < 0.05 ~ "*",
-    TRUE ~ ""
-  )) %>%
-  select(season, var, pval) %>%
-  spread(var, pval) %>%
-  rename(
-    par_total_pval = par_total,
-    rh_min_pval = rh_min,
-    temp_mean_pval = temp_mean
+```{r fig-5, evaluate = FALSE, fig.width=7, fig.height=8}
+# #' Extract p-values for plotting, and put into data frame
+# #' with xvals specifying season and yvals specifying height 
+# #' to plot asterisks.
+# p_vals <-
+#   t_test_results %>%
+#   mutate(pval = case_when(
+#     pval < 0.0001 ~ "***",
+#     pval < 0.001 ~ "**",
+#     pval < 0.05 ~ "*",
+#     TRUE ~ ""
+#   )) %>%
+#   select(season, var, pval) %>%
+#   spread(var, pval) %>%
+#   rename(
+#     par_total_pval = par_total,
+#     rh_min_pval = rh_min,
+#     temp_mean_pval = temp_mean
+#   )
+# 
+# # Simple function to calculate maximum y-position for plotting asterisks.
+# # Should be just above the mean + error bar height (1 sd).
+# maxy <- function(x) {mean(x, na.rm = TRUE) + sd(x, na.rm = TRUE)}
+# 
+# pval_plot_data <- 
+#   paired_microclimate %>%
+#   select(-date, -month) %>%
+#   group_by(season, site) %>%
+#   summarize_all(
+#     maxy
+#   ) %>%
+#   group_by(season) %>%
+#   summarize_all(
+#     max
+#   ) %>%
+#   select(-site) %>%
+#   left_join(p_vals)
+# 
+# #' Format data for plotting: need columns to be mean and sd, by site and season.
+# paired_means <- 
+#   paired_microclimate %>%
+#   select(-date) %>%
+#   group_by(site, season) %>%
+#   summarize_all(
+#     list(~mean(., na.rm = TRUE),
+#          ~sd(., na.rm = TRUE))
+#   ) %>%
+#   mutate(
+#     season = factor(season, levels = c("winter", "spring", "summer", "fall"))
+#   )
+# 
+# #' Make plot including t-test on three microclimatic variables.
+# #' 
+# #' The final plot includes multiple subplots. Each is a ridge plot showing 
+# #' the distribution of one of the variables between the two sites, with the
+# #' results of a paired t-test showing if they are different or not.
+# #' 
+# #' Need to manually set x ranges to keep x-axis scale the same across plots
+# x_ranges <- tibble(var = selected_vars) %>%
+#   mutate(
+#     min = map_dbl(var, ~ pull(paired_microclimate, .) %>% min ),
+#     max = map_dbl(var, ~ pull(paired_microclimate, .) %>% max )
+#   )
+# 
+# # light blue / light green colorblind-friendly from colorbrewer
+# # paired_colors <- c("#a6cee3", "#b2df8a")
+# # OR dark blue / dark green colorblind-friendly from colorbrewer
+# # paired_colors <- c("#1f78b4", "#33a02c")
+# # OR dark grey / light grey for bw
+# paired_colors <- c("grey55", "grey90")
+# 
+# subplots <- list()
+# 
+# #' Make ridge plots by mapping across nested datasets.
+# plot_results <-
+#   nested_paired_microclimate %>%
+#   right_join(x_ranges) %>%
+#   mutate(min = min - 0.05 * min,
+#          max = max + 0.05 * max,
+#          plot = pmap(list(var, data, x_min = min, x_max = max), 
+#                      plot_ridges, 
+#                      y_var = "site"),
+#          pval = map2(var, data, run_t_test),
+#          pval = case_when(
+#            pval < 0.0001 ~ "***",
+#            pval < 0.001 ~ "**",
+#            pval < 0.05 ~ "*",
+#            TRUE ~ ""
+#          ),
+#          plot = map2(
+#            plot, pval, 
+#            ~ .x + annotate("text", Inf, Inf, 
+#                            label = .y, hjust = 1, vjust = 1))
+#          
+#          # If needed, add annotation for season to make sure subplots
+#          # are positioned correctly
+#          
+#          # ,plot = map2(
+#          #   plot, season,
+#          #   ~ .x + annotate("text", -Inf, Inf,
+#          #                   label = .y, hjust = -1, vjust = 1)),
+#          # season = as_factor(season),
+#          # season = fct_relevel(season, "winter", "spring", "summer", "fall")
+#          
+#   ) %>% 
+#   arrange(season, var)
+# 
+# subplots <- plot_results$plot
+# 
+# #' Manually some subplots to remove redundant labels, etc.
+# subplots[1:9] <- map(subplots[1:9], 
+#                      ~ . + theme(axis.title.x = element_blank(),
+#                                  axis.text.x = element_blank() ))
+# 
+# 
+# subplots[1:12] <- map(subplots[1:12], 
+#                       ~ . + scale_fill_manual(
+#                         values = paired_colors,
+#                         name="Site",
+#                         breaks=c("okutama", "uratakao"),
+#                         labels=c("Okutama", "Uratakao"))
+# )
+# 
+# subplots[[10]] <- subplots[[10]] +
+#   labs(x = expression(paste("DLI (", mol~m^-2~day^-1, ")", sep = "")) )
+# 
+# subplots[[11]] <- subplots[[11]] +
+#   labs(x = expression("Min. Rel. Humidity (%)"))
+# 
+# subplots[[12]] <- subplots[[12]] +
+#   labs(x = expression("Mean Temp. (°C)"))
+# 
+# #' Add titles.
+# subplots <- map(subplots, ~ . + labs(title = "") )
+# 
+# subplots[[2]] <- subplots[[2]] +
+#   labs(title = "Winter\n")
+# 
+# subplots[[5]] <- subplots[[5]] +
+#   labs(title = "Spring\n")
+# 
+# subplots[[8]] <- subplots[[8]] +
+#   labs(title = "Summer\n")
+# 
+# subplots[[11]] <- subplots[[11]] +
+#   labs(title = "Fall\n")
+# 
+# #' Extract legend
+# legend <- paired_microclimate %>%
+#   ggplot(aes(x = temp_mean, y = site, fill = site)) +
+#   geom_density_ridges() +
+#   theme(legend.position = "bottom") +
+#   scale_fill_manual(
+#     values = paired_colors,
+#     name="Site",
+#     breaks=c("okutama", "uratakao"),
+#     labels=c("Okutama", "Uratakao"))
+# 
+# legend <- get_legend(legend)
+# 
+# #' Combine and write out.
+# subplots[[1]] +  subplots[[2]] +  subplots[[3]] + 
+#   subplots[[4]] +  subplots[[5]] +  subplots[[6]] + 
+#   subplots[[7]] +  subplots[[8]] +  subplots[[9]] + 
+#   subplots[[10]] + subplots[[11]] + subplots[[12]] +
+#   plot_spacer() + legend + plot_spacer() + plot_layout(ncol = 3, heights = c(1, 1, 1, 1, 0.2))
+```
+
+```{r table-2}
+cols_numeric <- c("r.squared", "adj.r.squared", "sigma", "statistic", "logLik", "AIC", "BIC")
+
+linear_model_results %>%
+  select(-fits) %>%
+  arrange(factor(x_var, levels = order_vars)) %>%
+  mutate(
+    x_var = case_when(
+      x_var == "rh_min" ~    "Min. Rel. Humidity",
+      x_var == "par_total" ~ "DLI",
+      x_var == "temp_mean" ~ "Mean Temp."
+    ),
+    y_var = case_when(
+      y_var == "total_cover" ~ "Total cover",
+      y_var == "count_mean" ~ "Gemmae count",
+      y_var == "length_mean" ~ "Gemmae length"
+    )
+  ) %>%
+  group_by(x_var) %>%
+  gt() %>%
+  cols_move_to_start(vars(y_var)) %>%
+  cols_move(
+    columns = vars(df.residual),
+    after = vars(df)
+  ) %>% 
+  fmt_number(
+    columns = colnames(.) %in% cols_numeric
+  ) %>%
+  fmt_number(
+    columns = vars(p.value),
+    decimals = 3
+  ) %>%
+  fmt_number(
+    columns = vars(deviance),
+    decimals = 0
+  ) %>%
+  cols_label(
+    y_var = "",
+    r.squared = "r-squared",
+    adj.r.squared = "adj. r-squared",
+    statistic = "t",
+    p.value = "p",
+    df.residual = "df residual"
+  ) %>%
+  tab_style(
+    style = cells_styles(
+      bkgd_color = "white"),
+    locations = cells_data(
+      columns = everything()
+    )
+  ) %>%
+  tab_header(
+    title = "Table 2"
   )
 
-# Simple function to calculate maximum y-position for plotting asterisks.
-# Should be just above the mean + error bar height (1 sd).
-maxy <- function(x) {mean(x, na.rm = TRUE) + sd(x, na.rm = TRUE)}
+```
 
-pval_plot_data <- 
-  paired_microclimate %>%
-  select(-date, -month) %>%
-  group_by(season, site) %>%
-  summarize_all(
-    maxy
-  ) %>%
-  group_by(season) %>%
-  summarize_all(
-    max
-  ) %>%
-  select(-site) %>%
-  left_join(p_vals)
+```{r fig-6, fig.width=7, fig.height=8}
+#' Loop through all combination of dep and indep variables and their models,
+#' and make plots for each.
+var_grid <- make_var_grid (c("rh_min", "par_total", "temp_mean"))
 
-#' Format data for plotting: need columns to be mean and sd, by site and season.
-paired_means <- 
-  paired_microclimate %>%
-  select(-date) %>%
-  group_by(site, season) %>%
-  summarize_all(
-    list(~mean(., na.rm = TRUE),
-         ~sd(., na.rm = TRUE))
-  ) %>%
-  mutate(
-    season = factor(season, levels = c("winter", "spring", "summer", "fall"))
+subplots <- 
+  pmap(var_grid, 
+       quick_plot_with_stats, 
+       plot_data = combined_monthly_morph, 
+       model_data = linear_model_results)
+
+#' Add axis titles
+#' 
+#' First, remove all default titles, then add formatted titles.
+subplots <- map(subplots, ~ . + labs(x = "", y = ""))
+subplots[[1]] <- subplots[[1]] + labs(y = expression("Total cover ("~cm^2~")"))
+subplots[[4]] <- subplots[[4]] + labs(y = "Gemmae count")
+subplots[[7]] <- subplots[[7]] + 
+  labs(
+    y = expression(paste("Gemmae len. (", mu, "m)", sep = "")),
+    x = expression(paste("DLI (", mol~m^-2~day^-1, ")", sep = ""))
+  )
+subplots[[8]] <- subplots[[8]] + labs(x = expression("Min. Rel. Humidity (%)"))
+subplots[[9]] <- subplots[[9]] + labs(x = expression("Mean Temp. (°C)"))
+
+#' Set amount of gap to close on left and top
+left_close <- -15
+top_close <- -10
+
+#' Close gaps on left only: 2,3
+subplots[2:3] <- subplots[2:3] %>%
+  map(~ . + theme(plot.margin = margin(0,0,0,left_close)) 
   )
 
-#' Make plot including t-test on three microclimatic variables.
-#' 
-#' The final plot includes multiple subplots. Each is a ridge plot showing 
-#' the distribution of one of the variables between the two sites, with the
-#' results of a paired t-test showing if they are different or not.
-#' 
-#' Need to manually set x ranges to keep x-axis scale the same across plots
-x_ranges <- tibble(var = selected_vars) %>%
-  mutate(
-    min = map_dbl(var, ~ pull(paired_microclimate, .) %>% min ),
-    max = map_dbl(var, ~ pull(paired_microclimate, .) %>% max )
+#' Close gaps on left side and top: 5,6,8,9
+subplots[c(5,6,8,9)] <- subplots[c(5,6,8,9)] %>%
+  map(~ . + theme(plot.margin = margin(top_close,0,0,left_close)) 
   )
 
-# light blue / light green colorblind-friendly from colorbrewer
-# paired_colors <- c("#a6cee3", "#b2df8a")
-# OR dark blue / dark green colorblind-friendly from colorbrewer
-# paired_colors <- c("#1f78b4", "#33a02c")
-# OR dark grey / light grey for bw
-paired_colors <- c("grey55", "grey90")
+#' Close gaps on top only: 4,7
+subplots[c(4,7)] <- subplots[c(4,7)] %>%
+  map(~ . + theme(plot.margin = margin(top_close,0,0,0)) 
+  )
+
+#' Add legend for middle bottom plot, combine, and write out.
+subplots[[8]] <- subplots[[8]] + 
+  theme(legend.position="bottom") + 
+  labs(color = "Month", shape = "Year")
+
+subplots[[1]] + subplots[[2]] + subplots[[3]] + 
+  subplots[[4]] + subplots[[5]] + subplots[[6]] + 
+  subplots[[7]] + subplots[[8]] + subplots[[9]] + 
+  plot_layout(ncol = 3)
+
+#' Black and white version if needed.
+# subplots <- 
+#   map(subplots,
+#       ~ . + scale_color_grey() 
+#   )
+# 
+# subplots[[1]] + subplots[[2]] + subplots[[3]] + 
+#   subplots[[4]] + subplots[[5]] + subplots[[6]] + 
+#   subplots[[7]] + subplots[[8]] + subplots[[9]] + 
+#   plot_layout(ncol = 3)
+```
+
+```{r fig-7, fig.width=7, fig.height=4}
+#' Note there is one outlier with length 46 mm, 
+#' then the rest all less than 19 mm. We will remove this 
+#' before plotting and make a note in the figure caption.
+gameto_size_no_outlier <-
+  gameto_size %>%
+  # Change sex to uppercase for nicer plotting
+  mutate(sex = stringr::str_to_title(sex)) %>%
+  # Remove one outlier
+  filter(length < 40)
+
+#' Make jitter plots overlayed with box plots. Make two
+#' subplots: one for width and one for length, then combine.
+
+# Use colorbrewer palette: 3 diverging, color-blind friendly colors
+cols <- c("#1b9e77","#d95f02","#7570b3")
+
+# Try to make sure dots stay in the same place, but doesn't work
+set.seed(1000)
+
+# Color version
 
@@
 
-#' Make ridge plots by mapping across nested datasets.
-plot_results <-
-  nested_paired_microclimate %>%
-  right_join(x_ranges) %>%
-  mutate(min = min - 0.05 * min,
-         max = max + 0.05 * max,
-         plot = pmap(list(var, data, x_min = min, x_max = max), 
-                     plot_ridges, 
-                     y_var = "site"),
-         pval = map2(var, data, run_t_test),
-         pval = case_when(
-           pval < 0.0001 ~ "***",
-           pval < 0.001 ~ "**",
-           pval < 0.05 ~ "*",
-           TRUE ~ ""
-         ),
-         plot = map2(
-           plot, pval, 
-           ~ .x + annotate("text", Inf, Inf, 
-                           label = .y, hjust = 1, vjust = 1))
+subplots[["length"]] <- 
+  ggplot(gameto_size_no_outlier, aes(y = length, x = sex, color = sex)) +
+  labs(
+    subtitle = "a",
+    y = "Length (mm)",
+    x = "",
+    color = "")
 
-         # If needed, add annotation for season to make sure subplots
-         # are positioned correctly
+subplots[["width"]] <-
+  ggplot(gameto_size_no_outlier, aes(y = width, x = sex, color = sex)) +
+  labs(
+    subtitle = "b",
+    y = "Width (mm)",
+    x = "")
 
-         # ,plot = map2(
-         #   plot, season,
-         #   ~ .x + annotate("text", -Inf, Inf,
-         #                   label = .y, hjust = -1, vjust = 1)),
-         # season = as_factor(season),
-         # season = fct_relevel(season, "winter", "spring", "summer", "fall")
-
-  ) %>% 
-  arrange(season, var)
-
-subplots <- plot_results$plot
-
-#' Manually some subplots to remove redundant labels, etc.
-subplots[1:9] <- map(subplots[1:9], 
-                     ~ . + theme(axis.title.x = element_blank(),
-                                 axis.text.x = element_blank() ))
-
-subplots[1:12] <- map(subplots[1:12], 
-                      ~ . + scale_fill_manual(
-                        values = paired_colors,
-                        name="Site",
-                        breaks=c("okutama", "uratakao"),
-                        labels=c("Okutama", "Uratakao"))
+subplots <- 
+  map(subplots,
+      ~ . + 
+        geom_jitter(size = 3) +
+        geom_boxplot(
+          outlier.shape = NA,
+          fill = NA,
+          color = "dark grey",
+          alpha = 0.7
+        ) +
+        scale_color_manual(values = cols) +
+        theme(legend.position = "none")
   )
 
-subplots[[10]] <- subplots[[10]] +
-  labs(x = expression(paste("DLI (", mol~m^-2~day^-1, ")", sep = "")) )
+subplots[[1]] + subplots[[2]] + plot_layout(ncol = 2)
 
-subplots[[11]] <- subplots[[11]] +
-  labs(x = expression("Min. Rel. Humidity (%)"))
+# Black and white version
+# subplots <- 
+#   map(subplots,
+#       ~ . + 
+#         scale_color_grey() +
+#         theme(legend.position = "none")
+#   )
 
-subplots[[12]] <- subplots[[12]] +
-  labs(x = expression("Mean Temp. (°C)"))
-
-#' Add titles.
-subplots <- map(subplots, ~ . + labs(title = "") )
-
-subplots[[2]] <- subplots[[2]] +
-  labs(title = "Winter\n")
-
-subplots[[5]] <- subplots[[5]] +
-  labs(title = "Spring\n")
-
-subplots[[8]] <- subplots[[8]] +
-  labs(title = "Summer\n")
-
-subplots[[11]] <- subplots[[11]] +
-  labs(title = "Fall\n")
-
-#' Extract legend
-legend <- paired_microclimate %>%
-  ggplot(aes(x = temp_mean, y = site, fill = site)) +
-  geom_density_ridges() +
-  theme(legend.position = "bottom") +
-  scale_fill_manual(
-    values = paired_colors,
-    name="Site",
-    breaks=c("okutama", "uratakao"),
-    labels=c("Okutama", "Uratakao"))
-
-legend <- get_legend(legend)
-
-#' Combine and write out.
-subplots[[1]] +  subplots[[2]] +  subplots[[3]] + 
-  subplots[[4]] +  subplots[[5]] +  subplots[[6]] + 
-  subplots[[7]] +  subplots[[8]] +  subplots[[9]] + 
-  subplots[[10]] + subplots[[11]] + subplots[[12]] +
-  plot_spacer() + legend + plot_spacer() + plot_layout(ncol = 3, heights = c(1, 1, 1, 1, 0.2))
 ```

### 2019-03-01 code/packages.R

@@ -1,33 +1,35 @@
 # Load the packages this workflow depends on
 
-# Install ferncollectR (Fern collection data)
-# This is updated frequently, so try to install 
-# the newest package first. The pacakge is a private 
-# github repo, so need have GITHUB_PAT defined in 
-# .Renviron to access.
-
-library(conflicted)
-library(assertthat)
 library(assertr)
+library(assertthat)
+library(broom)
+library(conflicted)
+library(cowplot)
 library(drake)
 library(glue)
+library(gt)
 library(here)
 library(jntools)
 library(lubridate)
 library(magrittr)
+library(patchwork)
 library(readxl)
+library(scales)
 library(tidyverse)
-library(broom)
+# library(ggridges)
 
-# Resolve conflicting functions
-collapse <- dplyr::collapse
-count <- dplyr::count
-expand <- tidyr::expand
-extract <- magrittr::extract
-filter <- dplyr::filter
-gather <- tidyr::gather
-intersect <- dplyr::intersect
-here <- here::here
-map <- purrr::map
-select <- dplyr::select
-setdiff <- dplyr::setdiff
+# Resolve conflicts
+col_factor <- conflict_prefer("col_factor", "readr")
+collapse   <- conflict_prefer("collapse", "dplyr")
+discard    <- conflict_prefer("discard", "purrr")
+expand     <- conflict_prefer("expand", "tidyr")
+extract    <- conflict_prefer("extract", "magrittr")
+filter     <- conflict_prefer("filter", "dplyr")
+gather     <- conflict_prefer("gather", "tidyr")
+ggsave     <- conflict_prefer("ggsave", "ggplot2")
+here       <- conflict_prefer("here", "here")
+intersect  <- conflict_prefer("intersect", "dplyr")
+lag        <- conflict_prefer("lag", "dplyr")
+Position   <- conflict_prefer("Position", "ggplot2")
+setdiff    <- conflict_prefer("setdiff", "dplyr")
+union      <- conflict_prefer("union", "dplyr")

### 2019-03-01 code/plan.R

@@ -118,6 +118,11 @@ data_cleaning_plan <- drake_plan (
     size_with_antheridia
   ),
   
+  #' ### Clean data
+  #' 
+  #' Select microclimate variables to use in downstream analyses
+  selected_vars = c("rh_min", "par_total", "temp_mean"),
+  
   #' Clean microclimate data
   okutama_microclimate = clean_microclimate(
     okutama_microclimate_raw),
@@ -146,7 +151,13 @@ data_cleaning_plan <- drake_plan (
   combined_morph = combine_morph_data (
     cover_data = cover,
     gemmae_data = gemmae_data
-  )
+  ),
+  
+  #' Combine morphology and Okutama climate into single tibble
+  combined_monthly_morph = combine_okutama_data (
+    okutama_microclimate = okutama_microclimate, 
+    combined_morph = combined_morph,
+    selected_vars = selected_vars)
 
 )
 
@@ -194,7 +205,7 @@ daily_summary_plan <- drake_plan (
   # Make nested data frame for looping.
   nested_paired_microclimate = nest_microclimate(
     paired_microclimate,
-    selected_vars = c("rh_min", "par_total", "temp_mean")
+    selected_vars = selected_vars
   )
  
 )
@@ -204,8 +215,20 @@ analysis_plan <- drake_plan (
   t_test_results = compare_microclimate(
     paired_microclimate = paired_microclimate,
     nested_paired_microclimate = nested_paired_microclimate,
-    selected_vars = c("rh_min", "par_total", "temp_mean")
-    )
+    selected_vars = selected_vars
+    ),
+  
+  linear_model_results = make_climate_morph_lm(
+    selected_vars = selected_vars,
+    combined_monthly_morph = combined_monthly_morph
+  ),
+  
+  # Write out MS
+  ms = rmarkdown::render(
+    knitr_in("ms.Rmd"),
+    output_format = "html_document",
+    output_file = file_out("ms.html"),
+    quiet = TRUE) 
   
 )
 

### 2019-03-01 make.R

@@ -1,6 +1,6 @@
 # make.R 
 
-# Master script for running T_angustifrons analyses.
+# Master script for running Pleurorosiopsis analyses.
 # This project uses drake to manage workflows.
 # For more information about drake, see
 # https://ropensci.github.io/drake/
@@ -20,7 +20,6 @@ pkgconfig::set_config("drake::strings_in_dots" = "literals")
 
 # General functions
 source("code/functions.R")
-source("code/drake_functions.R")
 
 # Data processing
 source("code/plan.R")

### 2019-03-01 code/01_clean_data.R
Delete old scripts now that everything has been merged into drake plan

@@ -1,429 +1,0 @@
-#+ include=FALSE
-# Suppress chatty tidyverse messages when spinning
-knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
-
-#' ## Load and clean data
-#' 
-#' This script processes the raw data files in `data_raw/` and outputs
-#' the cleaned data as `.csv` files to `data/`.
-#' 
-#' ### Setup
-setwd(here::here())
-
-source("code/setup.R")
-
-#' ### Data sources
-#' 
-#' The raw data are in `.xlsx` files. Some of the data are pre-processed
-#' (mean and sd of count and length of gemmae, 30 min averages of PPFD). 
-#' 
-#' The files include some additional analsyes and other data that won't 
-#' be used here.
-#'
-#' ### Microclimate data
-#'
-#' Microclimate variables include PPFD (photon flux density, in μmol of 
-#' light per sq m per sec), rel. humidity (%), and temperature (°C) 
-#' measured once every 30 min. 
-#' 
-#' PPFD is 30 min-averages of values taken every 4 minutes
-#' (raw 4 minute values not included).
-#' 
-#' There are two sites, Okutama (site of the independent gametophyte
-#' colony) and Uratakao (site of the sporophyte population). Additional
-#' sites were also measured, but not included in this analysis
-#' because of too much missing data due to mechanical failures.
-#' 
-#' For Okutama, the raw data are in different columns
-#' in the `xlsx` file, so read in each separately.
-okutama_temp_raw <- read_excel(
-  "data_raw/datalogger_raw.xlsm",
-  sheet = 1,
-  range = "A2:B28683",
-  col_names = c("date_time", "temp"))
-
-okutama_rh_raw <- read_excel(
-  "data_raw/datalogger_raw.xlsm",
-  sheet = 1,
-  range = "C2:D28683",
-  col_names = c("date_time", "rh"))
-
-okutama_par_raw <- read_excel(
-  "data_raw/datalogger_raw.xlsm",
-  sheet = 1,
-  range = "F2:G28204",
-  col_names = c("date_time", "par"))
-
-okutama_microclimate_raw <- 
-  full_join(okutama_temp_raw, okutama_rh_raw) %>%
-  full_join(okutama_par_raw)
-
-okutama_microclimate_raw
-
-#' Tidy Okutama microclimate data.
-#' 
[…]

### 2019-03-01 code/02_plot_growth.R

@@ -1,159 +1,0 @@
-#+ include=FALSE
-# Suppress chatty tidyverse messages when spinning
-knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
-
-#' ## Plot change in growth over time (Fig. 3)
-#'
-#' This script plots changes in growth (total area, 
-#' number and length of gemmae) in a population of independent
-#' gametophytes at Okutama, Japan.
-#'
-#' ### Setup
-setwd(here::here())
-
-source("code/setup.R")
-
-#' ### Load pre-processed data
-combined_morph <- read_csv("data/combined_morph.csv")
-
-#' Set x-limits manually since the automatically set limits will vary
-#' when we subset the data.
-start_date <- combined_morph %>%
-  pull(date) %>%
-  min
-
-end_date <- combined_morph %>%
-  pull(date) %>%
-  max
-
-#' Make tibble of years for shading. This dataframe will be passed to `geom_rect()` when
-#' plotting.
-shading_dates <- 
-  tibble(date = seq.Date(from = start_date, to = end_date, by = "day")) %>%
-  mutate(year = year(date)) %>%
-  # shade by alternate years
-  filter(year %in% c(2009, 2011, 2013)) %>%
-  group_by(year) %>%
-  summarize(
-    year_start = min(date),
-    year_end = max(date)
-  )
-
-#' I've also made a very specific function for adding rectangles to show
-#' the year:
-
-shade_years
-
-#' Make list of subplots. Build each one at a time, then add common
-#' formatting. 
-subplots <- list()
-
-subplots[["cover"]] <- 
-  combined_morph %>%
-  select(starts_with("q"), date) %>%
-  filter(complete.cases(.)) %>%
-  gather(plot, area, -date) %>%
-  ggplot(aes(x = date, y = area, linetype = plot)) %>%
-  shade_years(shading_dates) +
-  geom_line() +
-  scale_linetype_manual(
-    name="Quadrat",
-    breaks=c("q_1", "q_2", "q_3", "q_4"),
-    labels=c("1", "2", "3", "4"),
-    values=c("solid", "dashed", "dotted", "dotdash")
-  ) +
-  labs(
-    x = "",
-    y = expression("Cover ("~cm^2~")"),
-    subtitle = "a"
-  ) +
-  theme(legend.position = "bottom")
-
[…]

### 2019-03-01 code/03_plot_microclimate.R

@@ -1,168 +1,0 @@
-#+ include=FALSE
-# Suppress chatty tidyverse messages when spinning
-knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
-
-#' ## Plot change in microclimate over time (Fig. 4)
-#'
-#' This script plots changes in microclimate over time
-#' at two *Pleurosoriopsis makinoi* sites in Japan: Okutama
-#' and Uratakao. Okutama is home to a population of
-#' independent gametophytes of *P. makinoi*. At Uratakao,
-#' sporophytes and gametophytes grow together.
-#'   
-#' ### Setup
-setwd(here::here())
-
-source("code/setup.R")
-
-#' ### Load pre-processed data.
-daily_microclimate <- read_csv("data/daily_microclimate.csv")
-
-#' Extract common start and end dates so x-axis are same across subplots.
-start_date <- daily_microclimate %>% pull(date) %>% min
-end_date <- daily_microclimate %>% pull(date) %>% max
-all_days <- tibble( 
-  date = seq(start_date, end_date, by = "day")
-)
-
-#' Make tibble of years for shading.
-shading_dates <- 
-  tibble(date = seq.Date(from = start_date, to = end_date, by = "day")) %>%
-  mutate(year = year(date)) %>%
-  # shade by odd years
-  filter(year %in% c(2009, 2011, 2013)) %>%
-  group_by(year) %>%
-  summarize(
-    year_start = min(date),
-    year_end = max(date)
-  )
-
-#' PAR differs enough between sites that automatically set y-axis limits
-#' would be different. Use the common min and max between them.
-min_par <- daily_microclimate %>% pull(par_total) %>% min
-max_par <- daily_microclimate %>% pull(par_total) %>% max
-
-#' Insert NAs into microclimate for each site so that missing days
-#' don't get connected by lines.
-daily_microclimate_okutama <- filter(daily_microclimate, site == "okutama") %>%
-  right_join(all_days)
-
-daily_microclimate_takao <- filter(daily_microclimate, site == "uratakao") %>%
-  right_join(all_days)
-
-daily_microclimate_with_na <- bind_rows(
-  daily_microclimate_okutama, daily_microclimate_takao)
-
[…]

### 2019-03-01 code/04_compare_microclimate.R

@@ -1,361 +1,0 @@
-#+ include=FALSE
-# Suppress chatty tidyverse messages when spinning
-knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
-
-#' ## Compare microclimate between sites (Fig. 5 and Table 1)
-#'
-#' This script compares in microclimate 
-#' at two *Pleurosoriopsis makinoi* sites in Japan: Okutama
-#' and Uratakao. Okutama is home to a population of
-#' independent gametophytes of *P. makinoi*. At Uratakao,
-#' sporophytes and gametophytes grow together.
-#'   
-#' ### Setup
-setwd(here::here())
-
-source("code/setup.R")
-
-#' ### Load pre-processed data.
-daily_microclimate <- read_csv("data/daily_microclimate.csv")
-
-#' ## Fig 5 and Table 1: Compare microclimate between sites
-#'
-#' Subset the data to days only including means from both sites so they can
-#' be compared.
-
-# First get vector of days present in both sites, then use this to filter.
-both_days <-
-  daily_microclimate %>%
-  select(date, site, rh_min) %>%
-  spread(site, rh_min) %>%
-  drop_na() %>%
-  pull(date)
-
-paired_microclimate <- daily_microclimate %>%
-  filter(date %in% both_days)
-
-paired_microclimate
-
-#' Make a tibble of nested dataframes by season. We will loop across these
-#' and calculate mean and sd of each selected variable, and do a paired
-#' t-test between sites. 
-#' 
-# First make a cross-product dataframe including all seasons and
-# selected variables.
-var_grid <-
-  list(
-    var = selected_vars,
-    season = c("winter", "spring", "summer", "fall")
-  ) %>% cross_df()
-
-var_grid
-
-# Use `var_grid` to build a tibble of nested dataframes by season.
-nested_paired_microclimate <-
-  paired_microclimate %>%
-  select(date, site, season, selected_vars) %>%
-  group_by(season) %>%
-  nest %>%
-  right_join(var_grid)
-
-nested_paired_microclimate
-
-#' Calculate mean and sd for selected microclimate variables during each season,
-#' then add these to results of t-test. Keep everything wide for outputting to
-#' Table 1 in the paper.
-means_by_site_wide <-
[…]

### 2019-03-01 code/05_analyze_growth_microclimate.R

@@ -1,279 +1,0 @@
-#+ include=FALSE
-# Suppress chatty tidyverse messages when spinning
-knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
-
-#' ## Analyze relationship between growth and microclimate (Fig. 6 and Table 2)
-#'
-#' This script analyzes the relationship between growth 
-#' of gametophytes of *Pleurosoriopsis makinoi* and
-#' microclimate at Okutama, Japan.
-#' 
-#' Specifically, it makes a linear model for each combination 
-#' of microclimate variables (daily min RH, daily total PAR, 
-#' daily mean temp) as the independent variable and 
-#' growth traits (cover, number and length of gemmae) 
-#' as dependent variable.
-#'   
-#' ### Setup
-setwd(here::here())
-
-source("code/setup.R")
-
-#' ### Load pre-processed data
-okutama_microclimate <- read_csv("data/okutama_microclimate.csv")
-combined_morph <- read_csv("data/combined_morph.csv")
-
-#' ### Combine growth and microclimate data
-#' 
-#' Growth traits were measured once per month. To match these to microclimate,
-#' calculate monthly mean and sd of the three selected microclimate variables.
-okutama_monthly <-
-  okutama_microclimate %>%
-  filter(complete.cases(.)) %>% # no missing values, so don't need na.rm
-  group_by(date) %>%
-  summarize(
-    temp_mean = mean(temp, na.rm = TRUE),
-    rh_min = min(rh, na.rm = TRUE),
-    par_total = sum(1800*par) / 1000000
-  ) %>%
-  ungroup %>%
-  mutate(
-    month = month(date),
-    year = year(date),
-    month_year = paste(year, month, sep = "-")) %>%
-  group_by(month_year) %>%
-  summarize_at(
-    selected_vars,
-    list(
-      ~ mean(., na.rm = TRUE),
-      ~ sd(., na.rm = TRUE)
-    )
-  ) %>%
-  # For clarity, drop extra "mean" from variable names.
-  # Remember that we're dealing with monthly means though.
-  rename(
-    rh_min = rh_min_mean,
-    temp_mean = temp_mean_mean,
-    par_total = par_total_mean
-  )
-
-# Some of the growth data were measured before the
-# dataloggers were set up to measure mircroclimate.
[…]

### 2019-03-01 code/06_plot_gameto_size.R

@@ -1,102 +1,0 @@
-#+ include=FALSE
-# Suppress chatty tidyverse messages when spinning
-knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
-
-#' ## Analyze distribution of gametophyte size (Fig. 6)
-#'
-#' This script analyzes the distribution of individual sizes
-#' (length and width) of gametophytes of *Pleurosoriopsis makinoi* 
-#' at Okutama, Japan.
-#' 
-#' There were not enough measurments to do proper statistics, but
-#' we can visualize the distribution.
-#' 
-#' ### Setup
-setwd(here::here())
-
-source("code/setup.R")
-
-#' ### Load pre-processed data, and clean up a bit more.
-#' 
-#' Note there is one outlier with length 46 mm, 
-#' then the rest all less than 19 mm. We will remove this 
-#' before plotting and make a note in the figure caption.
-gameto_size <- read_csv("data/gameto_size.csv")
-
-gameto_size %>% arrange(desc(length))
-
-gameto_size <-
-  gameto_size %>%
-  # Change sex to uppercase for nicer plotting
-  mutate(sex = stringr::str_to_title(sex)) %>%
-  # Remove one outlier
-  filter(length < 40)
-
-gameto_size
-
-#' Make jitter plots overlayed with box plots. Make two
-#' subplots: one for width and one for length, then combine.
-
-# Use colorbrewer palette: 3 diverging, color-blind friendly colors
-cols <- c("#1b9e77","#d95f02","#7570b3")
-
-# Try to make sure dots stay in the same place, but doesn't work
-set.seed(1000)
-
-# Color version
-
-subplots <- list()
-
-subplots[["length"]] <- 
-  ggplot(gameto_size, aes(y = length, x = sex, color = sex)) +
-  labs(
-    subtitle = "a",
-    y = "Length (mm)",
-    x = "",
-    color = "")
-
-subplots[["width"]] <-
-  ggplot(gameto_size, aes(y = width, x = sex, color = sex)) +
-  labs(
-    subtitle = "b",
-    y = "Width (mm)",
-    x = "")
-
-subplots <- 
-  map(subplots,
-      ~ . + 
-        geom_jitter(size = 3) +
-        geom_boxplot(
-          outlier.shape = NA,
-          fill = NA,
-          color = "dark grey",
-          alpha = 0.7
-        ) +
[…]

### 2019-03-01 code/run_all.R

@@ -1,16 +1,0 @@
-#' Run all analyses.
-
-setwd(here::here())
-
-scripts <- list.files("code", pattern = "0")
-
-scripts <- sort(scripts)
-
-purrr::walk(scripts, 
-  ~ rmarkdown::render(
-  input = glue::glue("code/{.}"),
-  output_file = glue::glue("{.}_{Sys.Date()}.html"),
-  output_dir = "results/",
-  knit_root_dir = here::here(),
-  clean = TRUE)
-)

### 2019-03-01 code/setup.R

@@ -1,35 +1,0 @@
-#' # setup.R
-#' This script executes code that needs to be run before running any
-#' other analyses.
-
-#' Load packages.
-library(conflicted)
-library(tidyverse)
-library(readxl)
-library(lubridate)
-library(cowplot)
-library(broom)
-library(ggridges)
-library(patchwork)
-library(gt)
-library(viridis)
-
-#' Load custom functions.
-source("code/functions.R")
-
-#' Resolve namespace conflicts.
-conflict_prefer("filter", "dplyr")
-conflict_prefer("ggsave", "ggplot2")
-conflict_prefer("intersect", "dplyr")
-conflict_prefer("lag", "dplyr")
-conflict_prefer("Position", "ggplot2")
-conflict_prefer("scale_discrete_manual", "ggplot2")
-conflict_prefer("setdiff", "dplyr")
-conflict_prefer("union", "dplyr")
-
-#' Select microclimate variables of interest for analysis.
-selected_vars <- c("rh_min", "par_total", "temp_mean")
-
-#' Set order of microclimate variables for printing.
-order_vars <- c("par_total", "rh_min", "temp_mean")
-

### 2019-03-01 ms.Rmd
Add text of submitted MS

@@
 ---
-title: Growth Dynamics of the Independent Gametophytes of *Pleurorosiopsis makinoi* (Polypodiaceae)
+title: "Growth Dynamics of the Independent Gametophytes of *Pleurorosiopsis makinoi* (Polypodiaceae)"
 output: html_document
@@
 
-```{r fig-3, fig.width=8, fig.height=7}
+Atsushi Ebihara^1^\*, Joel H. Nitta^1^\*\*, Yurika Matsumoto^2^, Yuri Fukazawa^2^, Marie Kurihara^2^, Hitomi Yokote^2^, Kaoru Sakuma^2^, Otowa Azakami^2^, Yumiko Hirayama1 and Ryoko Imaichi^2^
+
+^1^Department of Botany, National Museum of Nature and Science, Amakubo 4–1–1, Tsukuba, Ibaraki 305–0005, Japan
+
+^2^Department of Chemical and Biological Sciences, Japan Women’s University, 2-8-1 Mejirodai, Bunkyo-ku, Tokyo 112-8681, Japan
+
+\* E-mail: ebihara@kahaku.go.jp
+
+\*\* Current affiliation: Department of Botany, National Museum of Natural History, Smithsonian Institute, Washington, D.C., 20013, U.S.A.
+
+(Received \*\*\*\*\*; accepted \*\*\*\*\*)
+
+Abstract Despite their importance to the fern life cycle, little is known about the ecology of fern gametophytes due to their cryptic morphology and small size.
+This study documents growth dynamics over five years in a population of Pleurosiopsis makinoi gametophytes growing independently of sporophytes in Okutama-machi, Tokyo, Japan.
+Cover was measured in four 10 x 10 cm quadrats, and the number and size of gemmae born on gametophytes measured ca.
+every two months from March 2009 to March 2014.
+The population survived for the duration of the study, and no seasonality was observed in the cover measurements.
+Gemmae gradually grow throughout the spring and summer, and mostly detach in winter.
+Based on our preliminary comparison between the microenvironment of the independent gametophyte site and a site supporting sporophytes of the same species, a possible cause of suppression of sporophyte formation is low light conditions in winter due to planting of evergreen conifers.
+
+Key words: fern, gametophyte, gemma, microenvironment, sporophyte
+
+Ferns possess an unusual life cycle consisting of two morphologically distinct, free-living stages: the diploid sporophyte and the haploid gametophyte.
+Unlike seed plants, in which the gametophyte is nutritionally dependent upon the sporophyte, in ferns gametophytes may grow separately from sporophytes.
+The phenomenon of ‘independent gametophytes’ includes fern gametophyte populations in various degrees of independence from sporophytes (Pinson et al., 2017).
+Independent gametophyte species, lacking any known counterpart sporophyte, are rather extreme cases (Farrar, 1967, 1978, 1992).
+Conversely, facultatively independent species, which may or may not grow separately from sporophytes depending on environmental conditions, have been observed in many more species (Rumsey and Sheffield, 1996).
+Almost without exception, species capable of growing as independent gametophytes have noncordiform morphology with indeterminate growth (Ebihara et al., 2013).
+Facultatively independent fern gametophytes represent an understudied but useful group for understanding the ecological correlates of the transition from haploid to diploid phase in land plants.
+
+We recently conducted a survey of the fern gametophyte flora of Japan (Ebihara et al., 2013).
+During the survey, we located a population of Pleurosoriopsis makinoi (Maxim.
+ex Makino) Fomin (Polypodiaceae) gametophytes in Unazawa Valley, Okutama-machi, Tokyo, Japan growing without any sporophytes nearby as far as we searched --- its sporophyte population has evidently been observed in the valley, but is very small and rare (S.
+Fujimoto, personal communication).
+We therefore considered it an ‘independent gametophyte’ in the broad sense.
+Pleurosoriopsis makinoi is a small-leaved fern species often growing with long-creeping rhizomes among bryophyte mats on rocks distributed in East Asia (China, Japan, Korea and Russian Far East) (Zhang et al., 2013).
+It has ribbon-like gametophytes, and gemmae and sexual organs have been observed both in the wild and in cultivation from sown spores (Masuyama, 1975).
+ We conducted a continuous observational study on the wild independent gametophytes of P. makinoi in Unazawa Valley to answer the following questions: 
+1.	How stable is the size of independent gametophyte populations over long periods of time?
+2.	Do independent gametophytes produce sexual organs (archegonia and antheridia) or possess the ability to produce sporophytes?
+3.	Do the gemmae of independent gametophytes contribute to sustainment/expansion of populations?
+4.	How do environmental factors affect gametophyte growth?
+
+## Materials and Methods
+
+### Location and habitat
+
+The independent gametophyte site is located in Unazawa Valley, Okutama-machi, Tokyo, Japan, at ca.
+315 m elevation.
+It grows on a small north-facing bank along a river (Fig.
+1).
+The population is ca.
+1 x 5 m in size.
+
+### Measurement of cover
+
+Four 10 cm × 10 cm quadrats (q1, q2, q3 and q4) were designated on the rockface (Fig.
+1) where the independent gametophyte population grows to measure cover (growth area).
+As the ribbon-like gametophytes of Hymenophyllum barbatum (Bosch) Baker (Hymenophyllaceae) also grow on the rockface, we carefully situated the quadrats to include only P.
+makinoi.
+We were able to distinguish between gametophytes of the two species based on the results of Ebihara et al.
+(2013) and morphological observation --- unlike P.
+makinoi, gametophytes of H.
+barbatum do not produce gemmae.
+Each quadrat was photographed using a digital camera.
+The image files were printed out, and a 1 cm2 grid was overlayed on the photo with tracing paper.
+Cover scores for each 1 cm2 cell in the grid were estimated in four discrete levels ranging from zero to 1, i.e.
+1.0 cm2 (100% ≥ cover > 50%), 0.5 cm2 (50% ≥ cover > 25%), 0.25 cm2 (25% ≥ cover > 0%), 0 (completely absent).
+Cover was calculated as the sum of the scores for each cell.
+The measurement was performed ca.
+every two months by placing the quadrats in exactly the same positions, and was repeated 37 times from March 2009 to March 2014.
+
+### Gametophyte collection, observation and measurements
+
+Living gametophytes were collected from the peripheries of the quadrats, and were brought back to the laboratory in small zipped plastic bags with wet papers.
+Collections were made 38 times from March 2009 to March 2014.
+The gametophytes were washed using brushes under a stereomicroscope (Leica M420).
+The number of gemmae was counted and length of each gemma (Figs.
+2a, b) was measured under a light microscope (Olympus BX51) in ten individuals per collection, and mean of the number and the length of gemmae calculated.
+
+We conducted a survey for reproductive structures in 2015.
+We collected 90 individuals, measured their size and width, and checked for the presence of antheridia or archegonia.
+During the 2015 survey, no reproductive structures were observed, so we also inspected images of all vouchers of P.
+makinoi collected at this site for archegonia or antheridia.
+
+In this study, an individual gametophyte is defined as a single distinct thallus.
+
+### Measurement of microenvironment
+
+HOBO Micro Station Data Loggers H21-002 (Onset) with sensors S-TMB-M002 (for temperature), S-THB-M002 (for temperature and humidity) and S-LIA-M003 for photosynthetic light (PAR) were used to measure microenvironment.
+In Unazawa Valley, the logger was installed from September 22, 2012 to July 22, 2014.
+To compare the environment of the independent gametophyte habitat with that of conspecific sporophytes, another set of the equipment was installed at a site where sporophytes were growing.
+The sporophyte site was at Uratakaomachi, Hachioji-shi, Tokyo, at 220 m elevation (from August 20, 2013 to November 22, 2014).
+Temperature and humidity were recorded every 30 minutes.
+PAR was sampled every 4 minutes and the average value was recorded every 30 minutes.
+
+The data recorded by the loggers were collected by connecting the logger to a notebook computer (Windows 7) and downloading using BoxCar Pro 4.3 and HOBOWare Pro every two to four months.
+The collected data were summarized by site as the following values: 1) daily mean temperature, 2) daily minimum humidity, 3) daily light integral (DLI).
+Daily light integral was calculated by summing the total PAR measurments over each day.
+
+### Statistical analysis
+
+Comparisons of microclimate between the sporophyte site and gametophyte site were made using a paired t-test.
+To test for the influence of microclimate on total cover (sum of all four quadrats), mean gemmae length, and mean gemmae size, linear models were fit with each of these as the dependent variable and the microclimatic variables as the independent variable.
+All statistical analysis was done using R 3.5.1 (R Core Team 2018).
+Plots were made with the “ggplot2” package (Wickham 2016).
+Raw data and code to replicate all analyses and figures are available at https://github.com/joelnitta/pleurosoriopsis.
+
+## Results
+
+### Measurement of cover
+
+The cover of Pleurosoriopsis gametophytes measured in four quadrats over five years is shown in Fig.
+3a.
+Cover irregularly fluctuated lower than 40% of original size, but never reached zero.
+Cover in three of the four quadrats (q1, q2, and q4) decreased during the last two years of the survey (2012–2014).
+
+### Measurement of number and size of gemmae
+
+The mean number and length of gemmae per individual are summarized in Figs.
+3b and 3c.
+The highest number of gemmae tend to be observed in late summer to autumn and the lowest number tend to be observed in spring (Fig.
+3b).
+Longer gemmae are frequently observed during November to January (Fig.
+3c).
+
+### Measurement of microenvironment
+
+The measured temperature, humidity and DLI are summarized in Fig 4.
+At the gametophyte site, data collection was interrupted by technical malfunction during the following periods: March 8–22, 2014, May 4–6, 2014, May 11–July 13, 2014.
+
+The sporophyte site and gametophyte site differed significantly in nearly all measured microclimatic variables (Fig.
+5, Table 1).
+The gametophyte site had lower light levels, especially in fall (0.21 ± 0.06 mmol per day, vs.
+0.33 ± 0.12 at the sporophyte site; p < 0.001, paired t¬-test) and winter (0.36 ± 0.27 mmol per day, vs.
+1.51 ± 0.74; p < 0.001), and tended to be drier, particularly in spring (50 ± 22% min.
+rel.
+humidity, vs.
+69 ± 21%; p < 0.001) and winter (78 ± 18%, vs.
+91 ± 10%; p < 0.001).
+Differences in temperature were significant but small (all estimated differences < 1°C; Table 1).
+
+### Relationship between growth and climate
+
+No significant relationship was observed between gametophyte growth and microclimate except for length of gemmae, which was negatively related to temperature (linear model; r2 = 0.51, p = 0.013) (Fig.
+6, Table 2).
+
+### Reproductive structures
+
+None of the gametophytes surveyed in 2015 possesed any sexual organs (n = 90).
+When we searched all other images of specimens collected from the Okutama site, archegonia (Figs.
+2c, d) were observed in 32 individuals (Fig.
+7).
+Antheridia (Figs.
+2e, f) were much less common than archegonia—we only found them in four individuals.
+Females tended to be larger than asexuals (it was not possible to conduct statistical analysis of size differences between sexes due to the low number of individuals with antheridia and inconsistency in sampling season and effort).
+No bisexual individuals were observed.
+We did not observed any juvenile sporophytes on the gametophytes.
+
+## Discussion
+
+The gametophyte patches in each of the four quadrats survived through all five years of our study.
+Even though we observed change in cover over time in each quadrat, there seems to be no seasonality in this fluctuation (Fig.
+3a).
+In contrast, clear seasonal patterns were observed in gemmae formation and growth (Figs.
+3b, c).
+Gemmae gradually grow and mostly detach in winter, the coldest and driest season.
+Considering the fact that no obvious new patches of gametophytes appeared during our study, the function of gemmae as asexual propagules was not demonstrated by our results.
+
+Although our measurement data of microenvironment are too short-termed to discuss the factors causing the independent gametophyte phenomenon, an interesting trend was observed in the DLI values—they are higher in the sporophyte site than the gametophyte site, particularly in the fall and winter (Fig.
+5, Table 1).
+The difference is undoubtedly caused by the different tree species covering the P.
+makinoi habitats—summer-green broad-leaved trees at the sporophyte site (Figs.
+8a, b) versus evergreen conifer [Chamaecyparis obtusa (Siebold et Zucc.) Endl.] plantation in Unazawa Valley, Okutama, the independent gametophyte site (Figs.
+8c, d).
+It is highly likely that sporophytes of P.
+makinoi, which have winter-green life history, are probably adapted to the summer-green forest floor environment and are not able to grow and survive on the evergreen forest floor due to a shortage of light.
+The stable presence of the gametophyte patches in Okutama suggests that the environment is suitable for gametophytes, but not sporophytes.
+Since we only observed a very small number of individuals with gametangia and no juvenile sporophytes, it seems likely that suppression of sporophyte production occurs prior to fertilization, or shortly thereafter.
+Archegoniate (female) individuals tended to be larger than asexual individuals (Fig.
+7), suggesting that gametophytes may need to reach a size threshold before they are able to sexually reproduce.
+The low winter light conditions at Okutama may prevent gametophytes from reaching this size consistently and thereby inhibit fertilization.
+
+“Inter-generational ecological niche separation”, or different ecological preferences between sporophytes and gametophytes of the same species, has been reported in several ferns (Sato and Sakai, 1981, Rumsey and Sheffield 1996, Pinson et al., 2017).
+In most cases, the gametophyte tends to tolerate a wider range of environments than the conspecific sporophyte (Watkins et al., 2007).
+This seems also to be the case for P.
+makinoi with regards to light condition.
+A possible scenario for formation of the independent gametophyte population in Okutama is as follows: 1) natural vegetation of the area was summer-green forest preferred by both sporophytes and gametophytes of P.
+makinoi; 2) after evergreen conifer plantations were established in the mid-20th century, sporophyte growth in the plantation areas was hampered by light shortage in winter and eventually ceased; 3) gametophytes still survive by indeterminate growth and vegetative reproduction by gemmae.
+
+We could not obtain any evidence for past presence of sporophytes at the site, but we found a similar case in Kyoto City, Kyoto Prefecture, Japan.
+At that site as well, evergreen conifer plantations dominated the forests starting from almost the same era.
+At least six herbarium specimens of Pleurosoriopsis sporophytes collected in the mountains surrounding Kyoto from 1928 to 1958 deposited in KYO, TNS and TOFO imply that it was not a very rare species at that time.
+Nowadays it is quite difficult to find Pleurosoriopsis sporophytes in the same area (Y.
+Kazumi, personal communication), but independent gametophytes are occasionally found (A.
+Ebihara, unpublished data).
+Such gametophytes whose sporophyte formation has been suppressed by recent environmental change during the past 100 years may resume their sexual life cycles when or if the environment returns to its prior condition.
+Further studies are necessary to determine the conditions needed for sporophyte formation, in particular frequencies of sexual organ formation.
+
+#### Acknowledgements
+
+This study was supported by a Grant-in-Aid for Scientific Research (KAKENHI 16370046 to Ryoko Imaichi) from the Japan Society of the Promotion of Science.
+
+## References
+
+Ebihara, A., Yamaoka, A., Mizukami, N., Sakoda, A., Nitta, J.
+H.
+and Imaichi, R.
+2013.
+A survey of the fern gametophyte flora of Japan: frequent independent occurrences of noncordiform gametophytes.
+Amer.
+J.
+Bot.
+100: 735-743.
+
+Farrar, D. R. 1967. Gametophytes of four tropical fern genera reproducing independently of their sporophytes in the southern Appalachians. Science 155: 1266–1267.
+
+Farrar, D. R. 1978. Problems in the identity and origin of the Appalachian Vittaria gametophyte, a sporophyteless fern of the eastern United States. American Journal of Botany 65: 1–12.
+
+Farrar, D. R. 1992. Trichomanes intricatum: the independent Trichomanes gametophyte in the eastern United States. American Fern Journal 82: 68–74.
+
+Masuyama, S. 1975. Gametophyte of *Pleurosoriopsis makinoi* (Maxim.) Fomin. Journal of Japanese Botany 50: 105-114.
+
+Pinson, J.
+B., Chambers, S.
+M., Nitta, J.
+H., Kuo, L.-Y.
+and Sessa, E.
+B.
+2017.
+The separation of generations: biology and biogeography of long-lived sporophyteless fern gametophytes.
+International Journal of Plant Sciences 178: 1-18.
+
+R Core Team. 2018. R: A language and environment for statistical computing. R Foundation for Statistical Computing, Vienna, Austria. https://www.R-project.org/.
+
+Rumsey, F.
+J.
+and Sheffield, E.
+1996.
+Inter-generational ecological niche separation and the 'independent gametophyte' phenomenon.
+In: Camus, J.
+M., Gibby, M.
+and Johns, R.
+J.
+(eds.), Pteridology in Perspective, pp.
+563-570.
+Royal Botanic Gardens, Kew, London.
+
+Sato, T. and Sakai, A. 1981. Cold tolerance of gametophytes and sporophytes of some cool temperate ferns native to Hokkaido. Canadian Journal of Botany 59: 604-608.
+
+Watkins Jr, J. E., Mack, M. K., and Mulkey, S. S. 2007. Gametophyte ecology and demography of epiphytic and terrestrial tropical ferns. American Journal of Botany 94: 701-708.
+
+Wickham, H. 2016. ggplot2: Elegant Graphics for Data Analysis. Springer-Verlag New York. https://ggplot2.tidyverse.org/.
+
+Zhang, X.
+C., Lu, S.
+G., Lin, Y.
+X., Qi, X.
+P., Moore, S., Xing, F.
+W, Wang, F.
+G., Hovenkamp, P.
+H., Gilbert, M.
+G., Nooteboom, H.
+P., Parris, B.
+S., Haufler, C., Kato, M.
+and Smith, A.
+R.
+2013.
+Polypodiaceae.
+In: Wu, Z.
+Y., Raven, P.
+H.
+and Hong, D.
+Y.
+(eds.), Flora of China, Vol.
+2-3 (Pteridophytes), pp.
+758-850.
+Science Press, Beijing, Missouri Botanical Garden Press, St.
+Louis.
+
+## Figures
+
+```{r fig-1, fig.width=7, fig.height=8, fig.cap = "Fig.
+1.
+The quadrats for measurement of gametophyte cover of Pleurosoriopsis makinoi in Unazawa Valley, Okutama-machi, Tokyo, Japan.
+a.
+A full view of the site with four quadrat positions (q1, q2, q3 and q4).
+b.
+A 10 cm x 10 cm quadrat frame."}
+
+magick::image_read_pdf(here("ms/Fig1_quadrats.pdf"))
+
+```
+
+```{r fig-2, fig.width=8, fig.height=7, fig.cap = "Fig.
+2.
+Gametophytes of Pleurosoriopsis makinoi collected in Unazawa Valley, Okutama-machi, Tokyo, Japan.
+a.
+A typical ribbon-like gametophyte with gemmae on the apical part (arrowhead), collected in August.
+b.
+Gemmae (arrowheads) on a gametophyte collected in November.
+c.
+A gametophyte with archegonia (arrowheads), collected in October.
+d.
+Archegonia.
+e.
+A gametophyte with antheridia (arrowhead) collected in December.
+f.
+An antheridium.
+Scale bars: 2 mm for a, 200 μm for b and d, 1 mm for c and e, 100 μm for f."}
+
+magick::image_read_pdf(here("ms/Fig2_gemma_sexualorgans.pdf"))
+
+```
+
+```{r fig-3, fig.width=8, fig.height=7, fig.cap="Fig.
+3.
+Change in growth of P.
+makinoi gametophytes observed from May 2009 to May 2014.
+a.
+Cover (cm2) of gametophytes in four 10 x 10 cm quadrats.
+b.
+Mean gemma length.
+c.
+Mean number of gemmae per individual.
+For b and c, n = 10 individuals per observation; grey bars show SD.
+Shaded bars show alternating years."}
 #' Set x-limits manually since the automatically set limits will vary
@@
 
-```{r fig-4, fig.width=8, fig.height=7}
+```{r fig-4, fig.width=8, fig.height=7, fig.cap = "Fig.
+4.
+Change in microclimate at site with P.
+makinoi independent gametophytes (Okutama) and sporophytes (Uratakao) from 2013 to 2014.
+a.
+Daily minimum relative humidity (%).
+b.
+Daily mean temperature (°C).
+c.
+Daily light integral (DLI) (mol m-2 day-1).
+Shaded bars show alternating years.
+Data collected at Okutama from September 22, 2012 to July 22, 2014 and at Uratakao from August 20, 2013 to November 22, 2014.
+Some data missing due to equipment malfunction (see Results)."}
 
@@
 
-```{r table-1}
-#' Format Table 1 and output to RTF using `gt` package.
-t_test_results %>%
-  arrange(factor(var, levels = order_vars)) %>%
-  mutate(season = stringr::str_to_title(season)) %>%
-  mutate(var = case_when(
-    var == "rh_min" ~    "Min. Rel. Humidity",
-    var == "par_total" ~ "DLI",
-    var == "temp_mean" ~ "Mean Temp."
-  )) %>%
-  group_by(season) %>%
-  gt() %>%
-  fmt_number(
-    columns = vars(estimate, statistic, conf.low, conf.high, 
-                   okutama_mean, okutama_sd, 
-                   uratakao_mean, uratakao_sd)
-  ) %>%
-  cols_move(
-    columns = vars(okutama_mean, okutama_sd),
-    after = vars(conf.high)
-  ) %>%
-  tab_spanner(
-    label = "Okutama",
-    columns = vars(okutama_mean, okutama_sd)
-  ) %>%
-  tab_spanner(
-    label = "Uratakao",
-    columns = vars(uratakao_mean, uratakao_sd)
-  ) %>%
-  cols_label(
-    var = "",
-    statistic = "t",
-    pval = "p",
-    parameter = "df",
-    conf.low = "Lower",
-    conf.high = "Upper",
-    okutama_mean = "Mean",
-    okutama_sd = "SD",
-    uratakao_mean = "Mean",
-    uratakao_sd = "SD"
-  ) %>%
-  tab_style(
-    style = cells_styles(
-      bkgd_color = "white"),
-    locations = cells_data(
-      columns = everything()
-    )
-  ) %>%
-  tab_header(
-    title = "Table 1"
-  )
-
-```
-
-```{r fig-5, evaluate = FALSE, fig.width=7, fig.height=8}
+```{r fig-5, evaluate = FALSE, fig.width=7, fig.height=8, fig.cap="Fig.
+5.
+Density plots showing difference between microclimate at P.
+makinoi independent gametophyte site (Okutama) and sporophyte site (Uratakao).
+Significant differences (paired t-test) shown by asterisks: ***, p < 0.001; **, p < 0.01; * p < 0.05.
+Y-axis is relative probability density; curves scaled so that total probability of each scales to 1.
+Vertical lines shown medians."}
 # #' Extract p-values for plotting, and put into data frame
@@
 
-```{r table-2}
-cols_numeric <- c("r.squared", "adj.r.squared", "sigma", "statistic", "logLik", "AIC", "BIC")
-
-linear_model_results %>%
-  select(-fits) %>%
-  arrange(factor(x_var, levels = order_vars)) %>%
-  mutate(
-    x_var = case_when(
-      x_var == "rh_min" ~    "Min. Rel. Humidity",
-      x_var == "par_total" ~ "DLI",
-      x_var == "temp_mean" ~ "Mean Temp."
-    ),
-    y_var = case_when(
-      y_var == "total_cover" ~ "Total cover",
-      y_var == "count_mean" ~ "Gemmae count",
-      y_var == "length_mean" ~ "Gemmae length"
-    )
-  ) %>%
-  group_by(x_var) %>%
-  gt() %>%
-  cols_move_to_start(vars(y_var)) %>%
-  cols_move(
-    columns = vars(df.residual),
-    after = vars(df)
-  ) %>% 
-  fmt_number(
-    columns = colnames(.) %in% cols_numeric
-  ) %>%
-  fmt_number(
-    columns = vars(p.value),
-    decimals = 3
-  ) %>%
-  fmt_number(
-    columns = vars(deviance),
-    decimals = 0
-  ) %>%
-  cols_label(
-    y_var = "",
-    r.squared = "r-squared",
-    adj.r.squared = "adj. r-squared",
-    statistic = "t",
-    p.value = "p",
-    df.residual = "df residual"
-  ) %>%
-  tab_style(
-    style = cells_styles(
-      bkgd_color = "white"),
-    locations = cells_data(
-      columns = everything()
-    )
-  ) %>%
-  tab_header(
-    title = "Table 2"
-  )
-
-```
-
-```{r fig-6, fig.width=7, fig.height=8}
+```{r fig-6, fig.width=7, fig.height=8, fig.cap="Fig.
+6.
+Relationships between growth of P.
+makinoi gametophytes (total cover, mean gemma number, and mean gemma length) and microclimate at the Okutama site.
+Microclimatic variables are monthly means of daily minimum relative humidity (%), daily mean temperature (°C), and daily light integral (DLI) (mol m-2 day-1).
+Grey bars show SD (except for total cover, which is a sum, not a mean).
+Significant relationships (linear model) shown with a line."}
 #' Loop through all combination of dep and indep variables and their models,
@@
 
-```{r fig-7, fig.width=7, fig.height=4}
+```{r fig-7, fig.width=7, fig.height=4, fig.cap = "Fig.
+7.
+Sizes of gametophytes with antheridia (male), archegonia (female), or no reproductive structures (asexual).
+a.
+Length (mm).
+b.
+Width (mm).
+Boxplots show medians and extend to the first and third quartile; whiskers extend to 1.5× the interquartile range.
+A single archegoniate (female) outlier with length of 46.6 mm was removed prior to plotting.
+Points randomly scattered within each x-axis category to reduce overlap."}
 #' Note there is one outlier with length 46 mm, 
@@
 ```
+
+```{r fig-8, fig.width=7, fig.height=8, fig.cap = "Fig.
+8.
+A comparison of vegetation of the independent gametophyte site with that of a sporophyte site of Pleurosoriopsis.
+a¬–b.
+Typical vegetation of sporophyte habitat covered with summer-green broad-leaved trees (Chichibu-shi, Saitama Pref., Japan).
+c–d.
+Vegetation of the independent gametophyte site in Unazawa Valley covered with evergreen cypress trees.
+a, c.
+in August, b, d.
+in December.
+"}
+
+magick::image_read_pdf(here("ms/Fig8_vegetation.pdf"))
+
+```
+
+## Tables
+
+```{r table-1, caption = "Results of paired t-test for difference in means between microclimatic variables (daily light integral [DLI], daily minimum relative humidity, and daily mean temperature) at two sites in Japan hosting independent gametophytes or sporophytes of Pleurosoriopsis makinoi (Okutama and Uratakao, respectively) by season."}
+#' Format Table 1 and output to RTF using `gt` package.
+t_test_results %>%
+  arrange(factor(var, levels = order_vars)) %>%
+  mutate(season = stringr::str_to_title(season)) %>%
+  mutate(var = case_when(
+    var == "rh_min" ~    "Min. Rel. Humidity",
+    var == "par_total" ~ "DLI",
+    var == "temp_mean" ~ "Mean Temp."
+  )) %>%
+  group_by(season) %>%
+  gt() %>%
+  fmt_number(
+    columns = vars(estimate, statistic, conf.low, conf.high, 
+                   okutama_mean, okutama_sd, 
+                   uratakao_mean, uratakao_sd)
+  ) %>%
+  cols_move(
+    columns = vars(okutama_mean, okutama_sd),
+    after = vars(conf.high)
+  ) %>%
+  tab_spanner(
+    label = "Okutama",
+    columns = vars(okutama_mean, okutama_sd)
+  ) %>%
+  tab_spanner(
+    label = "Uratakao",
+    columns = vars(uratakao_mean, uratakao_sd)
+  ) %>%
+  cols_label(
+    var = "",
+    statistic = "t",
+    pval = "p",
+    parameter = "df",
+    conf.low = "Lower",
+    conf.high = "Upper",
+    okutama_mean = "Mean",
+    okutama_sd = "SD",
+    uratakao_mean = "Mean",
+    uratakao_sd = "SD"
+  ) %>%
+  tab_style(
+    style = cells_styles(
+      bkgd_color = "white"),
+    locations = cells_data(
+      columns = everything()
+    )
+  ) %>%
+  tab_header(
+    title = "Table 1"
+  )
+
+```
+
+```{r table-2, caption = "Results of linear models testing for relationship between growth traits (total cover, count of gemmae, and length of gemmae) as dependent variable on microclimate (monthly averages of daily light integral [DLI], daily minimum relative humidity, and daily mean temperature) as independent variable in gametophytes of Pleurosoriopsis makinoi at Unazawa Valley, Okutama-machi, Tokyo, Japan.
+Results presented by independent variable."}
+cols_numeric <- c("r.squared", "adj.r.squared", "sigma", "statistic", "logLik", "AIC", "BIC")
+
+linear_model_results %>%
+  select(-fits) %>%
+  arrange(factor(x_var, levels = order_vars)) %>%
+  mutate(
+    x_var = case_when(
+      x_var == "rh_min" ~    "Min. Rel. Humidity",
+      x_var == "par_total" ~ "DLI",
+      x_var == "temp_mean" ~ "Mean Temp."
+    ),
+    y_var = case_when(
+      y_var == "total_cover" ~ "Total cover",
+      y_var == "count_mean" ~ "Gemmae count",
+      y_var == "length_mean" ~ "Gemmae length"
+    )
+  ) %>%
+  group_by(x_var) %>%
+  gt() %>%
+  cols_move_to_start(vars(y_var)) %>%
+  cols_move(
+    columns = vars(df.residual),
+    after = vars(df)
+  ) %>% 
+  fmt_number(
+    columns = colnames(.) %in% cols_numeric
+  ) %>%
+  fmt_number(
+    columns = vars(p.value),
+    decimals = 3
+  ) %>%
+  fmt_number(
+    columns = vars(deviance),
+    decimals = 0
+  ) %>%
+  cols_label(
+    y_var = "",
+    r.squared = "r-squared",
+    adj.r.squared = "adj. r-squared",
+    statistic = "t",
+    p.value = "p",
+    df.residual = "df residual"
+  ) %>%
+  tab_style(
+    style = cells_styles(
+      bkgd_color = "white"),
+    locations = cells_data(
+      columns = everything()
+    )
+  ) %>%
+  tab_header(
+    title = "Table 2"
+  )
+
+```

### 2019-05-04 ms.Rmd
Correct errors in proof
received from Ebihara 5-2-2019

Mostly spelling errors of Pleurosoriopsis

@@
 ---
-title: "Growth Dynamics of the Independent Gametophytes of *Pleurorosiopsis makinoi* (Polypodiaceae)"
+title: "Growth Dynamics of the Independent Gametophytes of *Pleurosoriopsis makinoi* (Polypodiaceae)"
 output: html_document
@@
 
-^1^Department of Botany, National Museum of Nature and Science, Amakubo 4–1–1, Tsukuba, Ibaraki 305–0005, Japan
+^1^Department of Botany, National Museum of Nature and Science, 4–1–1 Amakubo, Tsukuba, Ibaraki 305–0005, Japan
 
@@
 Abstract Despite their importance to the fern life cycle, little is known about the ecology of fern gametophytes due to their cryptic morphology and small size.
-This study documents growth dynamics over five years in a population of Pleurosiopsis makinoi gametophytes growing independently of sporophytes in Okutama-machi, Tokyo, Japan.
+This study documents growth dynamics over five years in a population of Pleurosoriopsis makinoi gametophytes growing independently of sporophytes in Okutama-machi, Tokyo, Japan.
 Cover was measured in four 10 x 10 cm quadrats, and the number and size of gemmae born on gametophytes measured ca.
@@
 3.
-Change in growth of P.
-makinoi gametophytes observed from May 2009 to May 2014.
+Change in growth of Pleurosoriopsis makinoi gametophytes observed from May 2009 to May 2014.
 a.
@@
 4.
-Change in microclimate at site with P.
-makinoi independent gametophytes (Okutama) and sporophytes (Uratakao) from 2013 to 2014.
+Change in microclimate at site with Pleurosoriopsis makinoi independent gametophytes (Okutama) and sporophytes (Uratakao) from 2013 to 2014.
 a.
@@
 5.
-Density plots showing difference between microclimate at P.
-makinoi independent gametophyte site (Okutama) and sporophyte site (Uratakao).
+Density plots showing difference between microclimate at Pleurosoriopsis makinoi independent gametophyte site (Okutama) and sporophyte site (Uratakao).
 Significant differences (paired t-test) shown by asterisks: ***, p < 0.001; **, p < 0.01; * p < 0.05.
@@
 6.
-Relationships between growth of P.
-makinoi gametophytes (total cover, mean gemma number, and mean gemma length) and microclimate at the Okutama site.
+Relationships between growth of Pleurosoriopsis makinoi gametophytes (total cover, mean gemma number, and mean gemma length) and microclimate at the Okutama site.
 Microclimatic variables are monthly means of daily minimum relative humidity (%), daily mean temperature (°C), and daily light integral (DLI) (mol m-2 day-1).

### 2019-05-06 install_packages.R
Use packrat with docker
in a single image to do analysis and render rmd

@@ -1,0 +1,112 @@
+#' # Installing packages to a Docker image with packrat
+#'
+#' ## How this works
+#'
+#' ### Setup
+#'
+#' Keep `Dockerfile` and `install_packages.R` in the root of the
+#' project directory.
+#'
+#' ### First step: make packrat/packrat.lock file
+#'
+#' Launch rocker/tidyverse with tag set to same version we will use in
+#' the Dockerfile and this directory mounted, and run this script.
+#'
+#' ````
+#' docker run --rm -e DISABLE_AUTH=true -it -v /Volumes/Transcend/Documents/Projects/pleurosoriopsis/:/home/rstudio/pleurosoriopsis rocker/verse:3.5.3 bash
+#' ````
+#'
+#' Inside the docker container first install any dependencies
+#' that aren't in rocker/verse. Then install R packages.
+#'
+#' ```
+#' apt-get update
+#' cd home/rstudio/pleurosoriopsis
+#' Rscript install_packages.R
+#' ```
+#'
+#' This will install current versions of all packages and deps
+#' called by library() below to that container,
+#' but the main reason is to write `packrat/packrat.lock`
+#'
+#' ### Second step: actually build the image
+#'
+#' Now we can use packrat.lock to restore (i.e., install) packages
+#' during the next (real) docker build, using the `Dockerfile`.
+#'
+#' ```
+#' docker build . -t joelnitta/pleurosoriopsis
+#' ```
+#'
+#' ### Third step: rinse, repeat
+#'
+#' Edit the packages below as needed to add new packages
+#' or update old ones (by installing), and repeat Steps 1 and 2.
+#'
+#' To only update new packages instead of running this entire script,
+#' install new packages IN THE CONTAINER using e.g.
+#' install.packages("visNetwork", lib = "/packrat/lib/x86_64-pc-linux-gnu/3.5.3", repos = "https://cran.rstudio.com/")
+#' (The `lib` argument ensures it installs to the packrat library).
+#' The run packrat::snapshot() again.
+#' (You should still add the package to the package list below so that the
+#' whole thing can be built from scratch if needed).
+
+################################################################################
+
+### Initialize packrat ###
+
+# Don't let packrat try to find
+# packages to install itself.
+
+install.packages("packrat", repos = "https://cran.rstudio.com/")
[…]

### 2019-05-06 code/plan.R
Don't use subplans

@@ -8,7 +8,9 @@
 #' The files include some additional analsyes and other data that won't 
 #' be used here.
 
-data_cleaning_plan <- drake_plan (
+plan <- drake_plan (
+  
+  # Load and clean data ----
   
   #' ### Microclimate data
   #'
@@ -157,11 +159,9 @@ data_cleaning_plan <- drake_plan (
   combined_monthly_morph = combine_okutama_data (
     okutama_microclimate = okutama_microclimate, 
     combined_morph = combined_morph,
-    selected_vars = selected_vars)
-
-)
-
-#' Calculate daily summary variables.
+    selected_vars = selected_vars),
+  
+# Calculate daily summary variables ----
 #' 
 #' In order to compare data across sites, we need to use daily values since
 #' the measurement times are slightly different between sites. 
@@ -185,8 +185,6 @@ data_cleaning_plan <- drake_plan (
 #' 
 #' sum of all μmol PAR in 30 min values in one day / 1,000,000 = total mol PAR for the day (DLI)
 
-daily_summary_plan <- drake_plan (
-  
   # Combine clean microclimata data from Okutama and Uratakao
   # into a single tibble with rows of daily values.
   daily_microclimate = calculate_daily_microclimate (
@@ -206,11 +204,9 @@ daily_summary_plan <- drake_plan (
   nested_paired_microclimate = nest_microclimate(
     paired_microclimate,
     selected_vars = selected_vars
-  )
- 
-)
+  ),
 
-analysis_plan <- drake_plan (
+# Analysis ----
   
   t_test_results = compare_microclimate(
     paired_microclimate = paired_microclimate,
@@ -233,5 +229,3 @@ analysis_plan <- drake_plan (
 )
 
 #### TO DO: start from 05_analyze_growth_microclimate
-
-plan <- bind_plans(data_cleaning_plan, daily_summary_plan, analysis_plan)

### 2019-05-06 code/functions.R
Assign col names to cover_raw

@@ -285,8 +285,8 @@ clean_microclimate <- function(raw_microclimate_data) {
 #' @return Tibble
 calculate_daily_microclimate <- function (site_1_data, site_1_name, site_2_data, site_2_name) {
   bind_rows(
-    mutate(okutama_microclimate, site = site_1_name),
-    mutate(takao_microclimate, site = site_2_name)
+    mutate(site_1_data, site = site_1_name),
+    mutate(site_2_data, site = site_2_name)
   ) %>%
     # only include days with data at both sites
     filter(complete.cases(.)) %>% 
@@ -317,7 +317,6 @@ calculate_daily_microclimate <- function (site_1_data, site_1_name, site_2_data,
 #' @return Tibble.
 clean_cover_data <- function(cover_raw) {
   cover_raw %>%
-    rename(date = "X__1", q_1 = `No. 1`, q_2 = `No. 2`, q_3 = `No. 3`, q_4 = `No. 4`) %>%
     mutate(date = lubridate::date(date)) %>%
     rowwise %>%
     mutate(total_cover = sum(q_1, q_2, q_3, q_4)) %>%

### 2019-05-06 code/plan.R

@@ -68,7 +68,10 @@ plan <- drake_plan (
   #' four 10 x 10 cm plots at the Okutama site. 
   cover_raw = read_excel(
     file_in("data_raw/pleurosoriopsis_data_figs.xlsx"),
-    sheet = 1),
+    sheet = 1, 
+    skip = 1,
+    col_names = c("date", "q_1", "q_2", "q_3", "q_4")
+    ),
   
   #' ### Gemmae
   #'

### 2019-05-06 ms.Rmd
Uncomment fig 5, load ggridges

@@
 Vertical lines shown medians."}
-# #' Extract p-values for plotting, and put into data frame
-# #' with xvals specifying season and yvals specifying height 
-# #' to plot asterisks.
-# p_vals <-
-#   t_test_results %>%
-#   mutate(pval = case_when(
-#     pval < 0.0001 ~ "***",
-#     pval < 0.001 ~ "**",
-#     pval < 0.05 ~ "*",
-#     TRUE ~ ""
-#   )) %>%
-#   select(season, var, pval) %>%
-#   spread(var, pval) %>%
-#   rename(
-#     par_total_pval = par_total,
-#     rh_min_pval = rh_min,
-#     temp_mean_pval = temp_mean
-#   )
-# 
-# # Simple function to calculate maximum y-position for plotting asterisks.
-# # Should be just above the mean + error bar height (1 sd).
-# maxy <- function(x) {mean(x, na.rm = TRUE) + sd(x, na.rm = TRUE)}
-# 
-# pval_plot_data <- 
-#   paired_microclimate %>%
-#   select(-date, -month) %>%
-#   group_by(season, site) %>%
-#   summarize_all(
-#     maxy
-#   ) %>%
-#   group_by(season) %>%
-#   summarize_all(
-#     max
-#   ) %>%
-#   select(-site) %>%
-#   left_join(p_vals)
-# 
-# #' Format data for plotting: need columns to be mean and sd, by site and season.
-# paired_means <- 
-#   paired_microclimate %>%
-#   select(-date) %>%
-#   group_by(site, season) %>%
-#   summarize_all(
-#     list(~mean(., na.rm = TRUE),
-#          ~sd(., na.rm = TRUE))
-#   ) %>%
-#   mutate(
-#     season = factor(season, levels = c("winter", "spring", "summer", "fall"))
-#   )
-# 
-# #' Make plot including t-test on three microclimatic variables.
-# #' 
-# #' The final plot includes multiple subplots. Each is a ridge plot showing 
-# #' the distribution of one of the variables between the two sites, with the
-# #' results of a paired t-test showing if they are different or not.
-# #' 
-# #' Need to manually set x ranges to keep x-axis scale the same across plots
-# x_ranges <- tibble(var = selected_vars) %>%
-#   mutate(
-#     min = map_dbl(var, ~ pull(paired_microclimate, .) %>% min ),
-#     max = map_dbl(var, ~ pull(paired_microclimate, .) %>% max )
-#   )
-# 
-# # light blue / light green colorblind-friendly from colorbrewer
-# # paired_colors <- c("#a6cee3", "#b2df8a")
-# # OR dark blue / dark green colorblind-friendly from colorbrewer
-# # paired_colors <- c("#1f78b4", "#33a02c")
-# # OR dark grey / light grey for bw
-# paired_colors <- c("grey55", "grey90")
-# 
-# subplots <- list()
-# 
-# #' Make ridge plots by mapping across nested datasets.
-# plot_results <-
-#   nested_paired_microclimate %>%
-#   right_join(x_ranges) %>%
-#   mutate(min = min - 0.05 * min,
-#          max = max + 0.05 * max,
-#          plot = pmap(list(var, data, x_min = min, x_max = max), 
-#                      plot_ridges, 
-#                      y_var = "site"),
-#          pval = map2(var, data, run_t_test),
-#          pval = case_when(
-#            pval < 0.0001 ~ "***",
-#            pval < 0.001 ~ "**",
-#            pval < 0.05 ~ "*",
-#            TRUE ~ ""
-#          ),
-#          plot = map2(
-#            plot, pval, 
-#            ~ .x + annotate("text", Inf, Inf, 
-#                            label = .y, hjust = 1, vjust = 1))
-#          
-#          # If needed, add annotation for season to make sure subplots
-#          # are positioned correctly
-#          
-#          # ,plot = map2(
-#          #   plot, season,
-#          #   ~ .x + annotate("text", -Inf, Inf,
-#          #                   label = .y, hjust = -1, vjust = 1)),
-#          # season = as_factor(season),
-#          # season = fct_relevel(season, "winter", "spring", "summer", "fall")
-#          
-#   ) %>% 
-#   arrange(season, var)
-# 
-# subplots <- plot_results$plot
-# 
-# #' Manually some subplots to remove redundant labels, etc.
-# subplots[1:9] <- map(subplots[1:9], 
-#                      ~ . + theme(axis.title.x = element_blank(),
-#                                  axis.text.x = element_blank() ))
-# 
-# 
-# subplots[1:12] <- map(subplots[1:12], 
-#                       ~ . + scale_fill_manual(
-#                         values = paired_colors,
-#                         name="Site",
-#                         breaks=c("okutama", "uratakao"),
-#                         labels=c("Okutama", "Uratakao"))
-# )
-# 
-# subplots[[10]] <- subplots[[10]] +
-#   labs(x = expression(paste("DLI (", mol~m^-2~day^-1, ")", sep = "")) )
-# 
-# subplots[[11]] <- subplots[[11]] +
-#   labs(x = expression("Min. Rel. Humidity (%)"))
-# 
-# subplots[[12]] <- subplots[[12]] +
-#   labs(x = expression("Mean Temp. (°C)"))
-# 
-# #' Add titles.
-# subplots <- map(subplots, ~ . + labs(title = "") )
-# 
-# subplots[[2]] <- subplots[[2]] +
-#   labs(title = "Winter\n")
-# 
-# subplots[[5]] <- subplots[[5]] +
-#   labs(title = "Spring\n")
-# 
-# subplots[[8]] <- subplots[[8]] +
-#   labs(title = "Summer\n")
-# 
-# subplots[[11]] <- subplots[[11]] +
-#   labs(title = "Fall\n")
-# 
-# #' Extract legend
-# legend <- paired_microclimate %>%
-#   ggplot(aes(x = temp_mean, y = site, fill = site)) +
-#   geom_density_ridges() +
-#   theme(legend.position = "bottom") +
-#   scale_fill_manual(
-#     values = paired_colors,
-#     name="Site",
-#     breaks=c("okutama", "uratakao"),
-#     labels=c("Okutama", "Uratakao"))
-# 
-# legend <- get_legend(legend)
-# 
-# #' Combine and write out.
-# subplots[[1]] +  subplots[[2]] +  subplots[[3]] + 
-#   subplots[[4]] +  subplots[[5]] +  subplots[[6]] + 
-#   subplots[[7]] +  subplots[[8]] +  subplots[[9]] + 
-#   subplots[[10]] + subplots[[11]] + subplots[[12]] +
-#   plot_spacer() + legend + plot_spacer() + plot_layout(ncol = 3, heights = c(1, 1, 1, 1, 0.2))
+#' Extract p-values for plotting, and put into data frame
+#' with xvals specifying season and yvals specifying height
+#' to plot asterisks.
+p_vals <-
+  t_test_results %>%
+  mutate(pval = case_when(
+    pval < 0.0001 ~ "***",
+    pval < 0.001 ~ "**",
+    pval < 0.05 ~ "*",
+    TRUE ~ ""
+  )) %>%
+  select(season, var, pval) %>%
+  spread(var, pval) %>%
+  rename(
+    par_total_pval = par_total,
+    rh_min_pval = rh_min,
+    temp_mean_pval = temp_mean
+  )
+
+# Simple function to calculate maximum y-position for plotting asterisks.
+# Should be just above the mean + error bar height (1 sd).
+maxy <- function(x) {mean(x, na.rm = TRUE) + sd(x, na.rm = TRUE)}
+
+pval_plot_data <-
+  paired_microclimate %>%
+  select(-date, -month) %>%
+  group_by(season, site) %>%
+  summarize_all(
+    maxy
+  ) %>%
+  group_by(season) %>%
+  summarize_all(
+    max
+  ) %>%
+  select(-site) %>%
+  left_join(p_vals)
+
+#' Format data for plotting: need columns to be mean and sd, by site and season.
+paired_means <-
+  paired_microclimate %>%
+  select(-date) %>%
+  group_by(site, season) %>%
+  summarize_all(
+    list(~mean(., na.rm = TRUE),
+         ~sd(., na.rm = TRUE))
+  ) %>%
+  mutate(
+    season = factor(season, levels = c("winter", "spring", "summer", "fall"))
+  )
+
+#' Make plot including t-test on three microclimatic variables.
+#'
+#' The final plot includes multiple subplots. Each is a ridge plot showing
+#' the distribution of one of the variables between the two sites, with the
+#' results of a paired t-test showing if they are different or not.
+#'
+#' Need to manually set x ranges to keep x-axis scale the same across plots
+x_ranges <- tibble(var = selected_vars) %>%
+  mutate(
+    min = map_dbl(var, ~ pull(paired_microclimate, .) %>% min ),
+    max = map_dbl(var, ~ pull(paired_microclimate, .) %>% max )
+  )
+
+# light blue / light green colorblind-friendly from colorbrewer
+# paired_colors <- c("#a6cee3", "#b2df8a")
+# OR dark blue / dark green colorblind-friendly from colorbrewer
+# paired_colors <- c("#1f78b4", "#33a02c")
+# OR dark grey / light grey for bw
+paired_colors <- c("grey55", "grey90")
+
+subplots <- list()
+
+#' Make ridge plots by mapping across nested datasets.
+plot_results <-
+  nested_paired_microclimate %>%
+  right_join(x_ranges) %>%
+  mutate(min = min - 0.05 * min,
+         max = max + 0.05 * max,
+         plot = pmap(list(var, data, x_min = min, x_max = max),
+                     plot_ridges,
+                     y_var = "site"),
+         pval = map2(var, data, run_t_test),
+         pval = case_when(
+           pval < 0.0001 ~ "***",
+           pval < 0.001 ~ "**",
+           pval < 0.05 ~ "*",
+           TRUE ~ ""
+         ),
+         plot = map2(
+           plot, pval,
+           ~ .x + annotate("text", Inf, Inf,
+                           label = .y, hjust = 1, vjust = 1))
+
+         # If needed, add annotation for season to make sure subplots
+         # are positioned correctly
+
+         # ,plot = map2(
+         #   plot, season,
+         #   ~ .x + annotate("text", -Inf, Inf,
+         #                   label = .y, hjust = -1, vjust = 1)),
+         # season = as_factor(season),
+         # season = fct_relevel(season, "winter", "spring", "summer", "fall")
+
+  ) %>%
+  arrange(season, var)
+
+subplots <- plot_results$plot
+
+#' Manually some subplots to remove redundant labels, etc.
+subplots[1:9] <- map(subplots[1:9],
+                     ~ . + theme(axis.title.x = element_blank(),
+                                 axis.text.x = element_blank() ))
+
+subplots[1:12] <- map(subplots[1:12],
+                      ~ . + scale_fill_manual(
+                        values = paired_colors,
+                        name="Site",
+                        breaks=c("okutama", "uratakao"),
+                        labels=c("Okutama", "Uratakao"))
+)
+
+subplots[[10]] <- subplots[[10]] +
+  labs(x = expression(paste("DLI (", mol~m^-2~day^-1, ")", sep = "")) )
+
+subplots[[11]] <- subplots[[11]] +
+  labs(x = expression("Min. Rel. Humidity (%)"))
+
+subplots[[12]] <- subplots[[12]] +
+  labs(x = expression("Mean Temp. (°C)"))
+
+#' Add titles.
+subplots <- map(subplots, ~ . + labs(title = "") )
+
+subplots[[2]] <- subplots[[2]] +
+  labs(title = "Winter\n")
+
+subplots[[5]] <- subplots[[5]] +
+  labs(title = "Spring\n")
+
+subplots[[8]] <- subplots[[8]] +
+  labs(title = "Summer\n")
+
+subplots[[11]] <- subplots[[11]] +
+  labs(title = "Fall\n")
+
+#' Extract legend
+legend <- paired_microclimate %>%
+  ggplot(aes(x = temp_mean, y = site, fill = site)) +
+  geom_density_ridges() +
+  theme(legend.position = "bottom") +
+  scale_fill_manual(
+    values = paired_colors,
+    name="Site",
+    breaks=c("okutama", "uratakao"),
+    labels=c("Okutama", "Uratakao"))
+
+legend <- get_legend(legend)
+
+#' Combine and write out.
+subplots[[1]] +  subplots[[2]] +  subplots[[3]] +
+  subplots[[4]] +  subplots[[5]] +  subplots[[6]] +
+  subplots[[7]] +  subplots[[8]] +  subplots[[9]] +
+  subplots[[10]] + subplots[[11]] + subplots[[12]] +
+  plot_spacer() + legend + plot_spacer() + plot_layout(ncol = 3, heights = c(1, 1, 1, 1, 0.2))
 ```

### 2019-05-06 code/packages.R

@@ -6,6 +6,7 @@ library(broom)
 library(conflicted)
 library(cowplot)
 library(drake)
+library(ggridges)
 library(glue)
 library(gt)
 library(here)

### 2019-05-06 code/plan.R

@@ -230,5 +230,3 @@ plan <- drake_plan (
     quiet = TRUE) 
   
 )
-
-#### TO DO: start from 05_analyze_growth_microclimate

### 2019-05-07 ms.Rmd
Fix season as factor
was getting converted back to character

@@
          ~sd(., na.rm = TRUE))
-  ) %>%
-  mutate(
-    season = factor(season, levels = c("winter", "spring", "summer", "fall"))
   )
@@
   ) %>%
+  mutate(
+    season = factor(season, levels = c("winter", "spring", "summer", "fall"))
+  ) %>%
   arrange(season, var)
@@
 #' Combine and write out.
-subplots[[1]] +  subplots[[2]] +  subplots[[3]] +
-  subplots[[4]] +  subplots[[5]] +  subplots[[6]] +
-  subplots[[7]] +  subplots[[8]] +  subplots[[9]] +
-  subplots[[10]] + subplots[[11]] + subplots[[12]] +
+subplots[[1]] +  subplots[[2]] +  subplots[[3]] + # Winter
+subplots[[4]] +  subplots[[5]] +  subplots[[6]] + # Spring
+subplots[[7]] +  subplots[[8]] +  subplots[[9]] + # Summer
+subplots[[10]] + subplots[[11]] + subplots[[12]] + # Fall
 plot_spacer() + legend + plot_spacer() + plot_layout(ncol = 3, heights = c(1, 1, 1, 1, 0.2))
+
+ggsave(glue::glue("fig-5_{Sys.Date()}.pdf"), width = 7, height = 8, units = "in")
 ```
@@
 6.
-Relationships between growth of Pleurosoriopsis makinoi gametophytes (total cover, mean gemma number, and mean gemma length) and microclimate at the Okutama site.
-Microclimatic variables are monthly means of daily minimum relative humidity (%), daily mean temperature (°C), and daily light integral (DLI) (mol m-2 day-1).
+Relationships between growth of *Pleurosoriopsis makinoi* gametophytes (total cover, mean gemma number, and mean gemma length) and microclimate at the Okutama site.
+Microclimatic variables are monthly means of daily light integral (DLI) (mol m^-2^ day^-1^), daily minimum relative humidity (%), and daily mean temperature (°C).
 Grey bars show SD (except for total cover, which is a sum, not a mean).

### 2019-05-07 ms.Rmd
Fix ms formatting

@@
 
-(Received \*\*\*\*\*; accepted \*\*\*\*\*)
-
-Abstract Despite their importance to the fern life cycle, little is known about the ecology of fern gametophytes due to their cryptic morphology and small size.
-This study documents growth dynamics over five years in a population of Pleurosoriopsis makinoi gametophytes growing independently of sporophytes in Okutama-machi, Tokyo, Japan.
-Cover was measured in four 10 x 10 cm quadrats, and the number and size of gemmae born on gametophytes measured ca.
+(Received 12 February 2019; accepted 27 March 2019)
+
+**Abstract** Despite their importance to the fern life cycle, little is known about the ecology of fern gametophytes due to their cryptic morphology and small size.
+This study documents growth dynamics over five years in a population of *Pleurosoriopsis makinoi* gametophytes growing independently of sporophytes in Okutama-machi, Tokyo, Japan.
+Cover was measured in four 10 × 10 cm quadrats, and the number and size of gemmae born on gametophytes measured ca.
 every two months from March 2009 to March 2014.
@@
 
-Key words: fern, gametophyte, gemma, microenvironment, sporophyte
+**Key words**: fern, gemma, microenvironment, sporophyte.
 
@@
 Unlike seed plants, in which the gametophyte is nutritionally dependent upon the sporophyte, in ferns gametophytes may grow separately from sporophytes.
-The phenomenon of ‘independent gametophytes’ includes fern gametophyte populations in various degrees of independence from sporophytes (Pinson et al., 2017).
+The phenomenon of ‘independent gametophytes’ includes fern gametophyte populations in various degrees of independence from sporophytes (Pinson *et al.*, 2017).
 Independent gametophyte species, lacking any known counterpart sporophyte, are rather extreme cases (Farrar, 1967, 1978, 1992).
 Conversely, facultatively independent species, which may or may not grow separately from sporophytes depending on environmental conditions, have been observed in many more species (Rumsey and Sheffield, 1996).
-Almost without exception, species capable of growing as independent gametophytes have noncordiform morphology with indeterminate growth (Ebihara et al., 2013).
+Almost without exception, species capable of growing as independent gametophytes have noncordiform morphology with indeterminate growth (Ebihara *et al.*, 2013).
 Facultatively independent fern gametophytes represent an understudied but useful group for understanding the ecological correlates of the transition from haploid to diploid phase in land plants.
 
-We recently conducted a survey of the fern gametophyte flora of Japan (Ebihara et al., 2013).
-During the survey, we located a population of Pleurosoriopsis makinoi (Maxim.
+We recently conducted a survey of the fern gametophyte flora of Japan (Ebihara *et al.*, 2013).
+During the survey, we located a population of *Pleurosoriopsis makinoi* (Maxim.
 ex Makino) Fomin (Polypodiaceae) gametophytes in Unazawa Valley, Okutama-machi, Tokyo, Japan growing without any sporophytes nearby as far as we searched --- its sporophyte population has evidently been observed in the valley, but is very small and rare (S.
@@
 We therefore considered it an ‘independent gametophyte’ in the broad sense.
-Pleurosoriopsis makinoi is a small-leaved fern species often growing with long-creeping rhizomes among bryophyte mats on rocks distributed in East Asia (China, Japan, Korea and Russian Far East) (Zhang et al., 2013).
+*Pleurosoriopsis makinoi* is a small-leaved fern species often growing with long-creeping rhizomes among bryophyte mats on rocks distributed in East Asia (China, Japan, Korea and Russian Far East) (Zhang *et al.*, 2013).
 It has ribbon-like gametophytes, and gemmae and sexual organs have been observed both in the wild and in cultivation from sown spores (Masuyama, 1975).
- We conducted a continuous observational study on the wild independent gametophytes of P. makinoi in Unazawa Valley to answer the following questions: 
+
+We conducted a continuous observational study on the wild independent gametophytes of *P. makinoi* in Unazawa Valley to answer the following questions: 
 1.	How stable is the size of independent gametophyte populations over long periods of time?
@@
 The population is ca.
-1 x 5 m in size.
+1 × 5 m in size.
 
@@
 1) where the independent gametophyte population grows to measure cover (growth area).
-As the ribbon-like gametophytes of Hymenophyllum barbatum (Bosch) Baker (Hymenophyllaceae) also grow on the rockface, we carefully situated the quadrats to include only P.
-makinoi.
-We were able to distinguish between gametophytes of the two species based on the results of Ebihara et al.
-(2013) and morphological observation --- unlike P.
-makinoi, gametophytes of H.
-barbatum do not produce gemmae.
+As the ribbon-like gametophytes of *Hymenophyllum barbatum* (Bosch) Baker (Hymenophyllaceae) also grow on the rockface, we carefully situated the quadrats to include only *P.
+makinoi*.
+We were able to distinguish between gametophytes of the two species based on the results of Ebihara *et al.* (2013) and morphological observation --- unlike *P.
+makinoi*, gametophytes of *H.
+barbatum* do not produce gemmae.
 Each quadrat was photographed using a digital camera.
-The image files were printed out, and a 1 cm2 grid was overlayed on the photo with tracing paper.
-Cover scores for each 1 cm2 cell in the grid were estimated in four discrete levels ranging from zero to 1, i.e.
-1.0 cm2 (100% ≥ cover > 50%), 0.5 cm2 (50% ≥ cover > 25%), 0.25 cm2 (25% ≥ cover > 0%), 0 (completely absent).
+The image files were printed out, and a 1 cm^2^ grid was overlayed on the photo with tracing paper.
+Cover scores for each 1 cm^2^ cell in the grid were estimated in four discrete levels ranging from zero to 1, i.e.
+1.0 cm^2^ (100% ≥ cover > 50%), 0.5 cm^2^ (50% ≥ cover > 25%), 0.25 cm^2^ (25% ≥ cover > 0%), 0 (completely absent).
 Cover was calculated as the sum of the scores for each cell.
@@
 We conducted a survey for reproductive structures in 2015.
-We collected 90 individuals, measured their size and width, and checked for the presence of antheridia or archegonia.
-During the 2015 survey, no reproductive structures were observed, so we also inspected images of all vouchers of P.
-makinoi collected at this site for archegonia or antheridia.
+We collected 90 individuals, measured their length and width, and checked for the presence of antheridia or archegonia.
+During the 2015 survey, no reproductive structures were observed, so we also inspected images of all vouchers of *P.
+makinoi* collected at this site for archegonia or antheridia.
 
@@
 
-Comparisons of microclimate between the sporophyte site and gametophyte site were made using a paired t-test.
-To test for the influence of microclimate on total cover (sum of all four quadrats), mean gemmae length, and mean gemmae size, linear models were fit with each of these as the dependent variable and the microclimatic variables as the independent variable.
+Comparisons of microclimate between the sporophyte site and gametophyte site were made using a paired *t*-test.
+To test for the influence of microclimate on total cover (sum of all four quadrats), mean gemmae length, and mean gemmae number, linear models were fit with each of these as the dependent variable and the microclimatic variables as the independent variable.
 All statistical analysis was done using R 3.5.1 (R Core Team 2018).
@@
 The gametophyte site had lower light levels, especially in fall (0.21 ± 0.06 mmol per day, vs.
-0.33 ± 0.12 at the sporophyte site; p < 0.001, paired t¬-test) and winter (0.36 ± 0.27 mmol per day, vs.
-1.51 ± 0.74; p < 0.001), and tended to be drier, particularly in spring (50 ± 22% min.
+0.33 ± 0.12 at the sporophyte site; *p* < 0.001, paired *t*-test) and winter (0.36 ± 0.27 mmol per day, vs.
+1.51 ± 0.74; *p* < 0.001), and tended to be drier, particularly in spring (50 ± 22% min.
 rel.
 humidity, vs.
-69 ± 21%; p < 0.001) and winter (78 ± 18%, vs.
-91 ± 10%; p < 0.001).
+69 ± 21%; *p* < 0.001) and winter (78 ± 18%, vs.
+91 ± 10%; *p* < 0.001).
 Differences in temperature were significant but small (all estimated differences < 1°C; Table 1).
@@
 
-No significant relationship was observed between gametophyte growth and microclimate except for length of gemmae, which was negatively related to temperature (linear model; r2 = 0.51, p = 0.013) (Fig.
+No significant relationship was observed between gametophyte growth and microclimate except for length of gemmae, which was negatively related to temperature (linear model; *r*^2^ = 0.51, *p* = 0.013) (Fig.
 6, Table 2).
@@
 
-None of the gametophytes surveyed in 2015 possesed any sexual organs (n = 90).
+None of the gametophytes surveyed in 2015 possesed any sexual organs (*n* = 90).
 When we searched all other images of specimens collected from the Okutama site, archegonia (Figs.
@@
 Antheridia (Figs.
-2e, f) were much less common than archegonia—we only found them in four individuals.
+2e, f) were much less common than archegonia --- we only found them in four individuals.
 Females tended to be larger than asexuals (it was not possible to conduct statistical analysis of size differences between sexes due to the low number of individuals with antheridia and inconsistency in sampling season and effort).
 No bisexual individuals were observed.
-We did not observed any juvenile sporophytes on the gametophytes.
+We did not observe any juvenile sporophytes on the gametophytes.
 
@@
 
-Although our measurement data of microenvironment are too short-termed to discuss the factors causing the independent gametophyte phenomenon, an interesting trend was observed in the DLI values—they are higher in the sporophyte site than the gametophyte site, particularly in the fall and winter (Fig.
+Although our measurement data of microenvironment are too short-termed to discuss the factors causing the independent gametophyte phenomenon, an interesting trend was observed in the DLI values --- they are higher in the sporophyte site than the gametophyte site, particularly in the fall and winter (Fig.
 5, Table 1).
-The difference is undoubtedly caused by the different tree species covering the P.
-makinoi habitats—summer-green broad-leaved trees at the sporophyte site (Figs.
+The difference is undoubtedly caused by the different tree species covering the *P.
+makinoi* habitats --- summer-green broad-leaved trees at the sporophyte site (Figs.
 8a, b) versus evergreen conifer [Chamaecyparis obtusa (Siebold et Zucc.) Endl.] plantation in Unazawa Valley, Okutama, the independent gametophyte site (Figs.
 8c, d).
-It is highly likely that sporophytes of P.
-makinoi, which have winter-green life history, are probably adapted to the summer-green forest floor environment and are not able to grow and survive on the evergreen forest floor due to a shortage of light.
+It is highly likely that sporophytes of *P.
+makinoi*, which have winter-green life history, are probably adapted to the summer-green forest floor environment and are not able to grow and survive on the evergreen forest floor due to a shortage of light.
 The stable presence of the gametophyte patches in Okutama suggests that the environment is suitable for gametophytes, but not sporophytes.
@@
 
-“Inter-generational ecological niche separation”, or different ecological preferences between sporophytes and gametophytes of the same species, has been reported in several ferns (Sato and Sakai, 1981, Rumsey and Sheffield 1996, Pinson et al., 2017).
-In most cases, the gametophyte tends to tolerate a wider range of environments than the conspecific sporophyte (Watkins et al., 2007).
-This seems also to be the case for P.
-makinoi with regards to light condition.
-A possible scenario for formation of the independent gametophyte population in Okutama is as follows: 1) natural vegetation of the area was summer-green forest preferred by both sporophytes and gametophytes of P.
-makinoi; 2) after evergreen conifer plantations were established in the mid-20th century, sporophyte growth in the plantation areas was hampered by light shortage in winter and eventually ceased; 3) gametophytes still survive by indeterminate growth and vegetative reproduction by gemmae.
+“Inter-generational ecological niche separation”, or different ecological preferences between sporophytes and gametophytes of the same species, has been reported in several ferns (Sato and Sakai, 1981, Rumsey and Sheffield 1996, Pinson *et al.*, 2017).
+In most cases, the gametophyte tends to tolerate a wider range of environments than the conspecific sporophyte (Watkins *et al.*, 2007).
+This seems also to be the case for *P.
+makinoi* with regards to light condition.
+A possible scenario for formation of the independent gametophyte population in Okutama is as follows: 1) natural vegetation of the area was summer-green forest preferred by both sporophytes and gametophytes of *P.
+makinoi*; 2) after evergreen conifer plantations were established in the mid-20th century, sporophyte growth in the plantation areas was hampered by light shortage in winter and eventually ceased; 3) gametophytes still survive by indeterminate growth and vegetative reproduction by gemmae.
 
@@
 At that site as well, evergreen conifer plantations dominated the forests starting from almost the same era.
-At least six herbarium specimens of Pleurosoriopsis sporophytes collected in the mountains surrounding Kyoto from 1928 to 1958 deposited in KYO, TNS and TOFO imply that it was not a very rare species at that time.
-Nowadays it is quite difficult to find Pleurosoriopsis sporophytes in the same area (Y.
+At least six herbarium specimens of *Pleurosoriopsis* sporophytes collected in the mountains surrounding Kyoto from 1928 to 1958 deposited in KYO, TNS and TOFO imply that it was not a very rare species at that time.
+Nowadays it is quite difficult to find *Pleurosoriopsis* sporophytes in the same area (Y.
 Kazumi, personal communication), but independent gametophytes are occasionally found (A.
@@
 
-Farrar, D. R. 1978. Problems in the identity and origin of the Appalachian Vittaria gametophyte, a sporophyteless fern of the eastern United States. American Journal of Botany 65: 1–12.
-
-Farrar, D. R. 1992. Trichomanes intricatum: the independent Trichomanes gametophyte in the eastern United States. American Fern Journal 82: 68–74.
+Farrar, D. R. 1978. Problems in the identity and origin of the Appalachian *Vittaria* gametophyte, a sporophyteless fern of the eastern United States. American Journal of Botany 65: 1–12.
+
+Farrar, D. R. 1992. *Trichomanes intricatum*: the independent *Trichomanes* gametophyte in the eastern United States. American Fern Journal 82: 68–74.
 
@@
 1.
-The quadrats for measurement of gametophyte cover of Pleurosoriopsis makinoi in Unazawa Valley, Okutama-machi, Tokyo, Japan.
+The quadrats for measurement of gametophyte cover of *Pleurosoriopsis makinoi* in Unazawa Valley, Okutama-machi, Tokyo, Japan.
 a.
@@
 b.
-A 10 cm x 10 cm quadrat frame."}
+A 10 cm × 10 cm quadrat frame."}
 
@@
 2.
-Gametophytes of Pleurosoriopsis makinoi collected in Unazawa Valley, Okutama-machi, Tokyo, Japan.
+Gametophytes of *Pleurosoriopsis makinoi* collected in Unazawa Valley, Okutama-machi, Tokyo, Japan.
 a.
@@
 3.
-Change in growth of Pleurosoriopsis makinoi gametophytes observed from May 2009 to May 2014.
+Change in growth of *Pleurosoriopsis makinoi* gametophytes observed from May 2009 to May 2014.
 a.
-Cover (cm2) of gametophytes in four 10 x 10 cm quadrats.
+Cover (cm^2^) of gametophytes in four 10 × 10 cm quadrats.
 b.
-Mean gemma length.
+Mean number of gemmae per individual.
 c.
-Mean number of gemmae per individual.
-For b and c, n = 10 individuals per observation; grey bars show SD.
+Mean gemma length (μm).
+For b and c, *n* = 10 individuals per observation; grey bars show SD.
 Shaded bars show alternating years."}
@@
 4.
-Change in microclimate at site with Pleurosoriopsis makinoi independent gametophytes (Okutama) and sporophytes (Uratakao) from 2013 to 2014.
-a.
+Change in microclimate at site with *Pleurosoriopsis makinoi* independent gametophytes (Okutama) and sporophytes (Uratakao) from 2013 to 2014.
+a--b.
+Daily light integral (DLI) (mol m^-2 day^-1).
+c--d.
 Daily minimum relative humidity (%).
-b.
+e--f.
 Daily mean temperature (°C).
-c.
-Daily light integral (DLI) (mol m-2 day-1).
 Shaded bars show alternating years.
@@
 5.
-Density plots showing difference between microclimate at Pleurosoriopsis makinoi independent gametophyte site (Okutama) and sporophyte site (Uratakao).
-Significant differences (paired t-test) shown by asterisks: ***, p < 0.001; **, p < 0.01; * p < 0.05.
+Density plots showing difference between microclimate at *Pleurosoriopsis makinoi* independent gametophyte site (Okutama) and sporophyte site (Uratakao).
+Significant differences (paired t-test) shown by asterisks: \\*\\*\\*, *p* < 0.001; \\*\\*, *p* < 0.01; \\*, *p* < 0.05.
 Y-axis is relative probability density; curves scaled so that total probability of each scales to 1.
@@
 A comparison of vegetation of the independent gametophyte site with that of a sporophyte site of Pleurosoriopsis.
-a¬–b.
+a--b.
 Typical vegetation of sporophyte habitat covered with summer-green broad-leaved trees (Chichibu-shi, Saitama Pref., Japan).
-c–d.
+c--d.
 Vegetation of the independent gametophyte site in Unazawa Valley covered with evergreen cypress trees.
@@
 
-```{r table-1, caption = "Results of paired t-test for difference in means between microclimatic variables (daily light integral [DLI], daily minimum relative humidity, and daily mean temperature) at two sites in Japan hosting independent gametophytes or sporophytes of Pleurosoriopsis makinoi (Okutama and Uratakao, respectively) by season."}
+Table 1.
+Results of paired *t*-test for difference in means between microclimatic variables (daily light integral [DLI], daily minimum relative humidity, and daily mean temperature) at two sites in Japan hosting independent gametophytes or sporophytes of *Pleurosoriopsis makinoi* (Okutama and Uratakao, respectively) by season.
+
+```{r table-1}
 #' Format Table 1 and output to RTF using `gt` package.
@@
 
-```{r table-2, caption = "Results of linear models testing for relationship between growth traits (total cover, count of gemmae, and length of gemmae) as dependent variable on microclimate (monthly averages of daily light integral [DLI], daily minimum relative humidity, and daily mean temperature) as independent variable in gametophytes of Pleurosoriopsis makinoi at Unazawa Valley, Okutama-machi, Tokyo, Japan.
-Results presented by independent variable."}
+Table 2.
+Results of linear models testing for relationship between growth traits (total cover, count of gemmae, and length of gemmae) as dependent variable on microclimate (monthly averages of daily light integral [DLI], daily minimum relative humidity, and daily mean temperature) as independent variable in gametophytes of *Pleurosoriopsis makinoi* at Unazawa Valley, Okutama-machi, Tokyo, Japan.
+Results presented by independent variable.
+
+```{r table-2}
 cols_numeric <- c("r.squared", "adj.r.squared", "sigma", "statistic", "logLik", "AIC", "BIC")

### 2019-06-20 install_packages.R
Update R in container to 3.6.0

@@ -1,57 +1,7 @@
-#' # Installing packages to a Docker image with packrat
-#'
-#' ## How this works
-#'
-#' ### Setup
-#'
-#' Keep `Dockerfile` and `install_packages.R` in the root of the
-#' project directory.
-#'
-#' ### First step: make packrat/packrat.lock file
-#'
-#' Launch rocker/tidyverse with tag set to same version we will use in
-#' the Dockerfile and this directory mounted, and run this script.
-#'
-#' ````
-#' docker run --rm -e DISABLE_AUTH=true -it -v /Volumes/Transcend/Documents/Projects/pleurosoriopsis/:/home/rstudio/pleurosoriopsis rocker/verse:3.5.3 bash
-#' ````
-#'
-#' Inside the docker container first install any dependencies
-#' that aren't in rocker/verse. Then install R packages.
-#'
-#' ```
-#' apt-get update
-#' cd home/rstudio/pleurosoriopsis
-#' Rscript install_packages.R
-#' ```
-#'
-#' This will install current versions of all packages and deps
-#' called by library() below to that container,
-#' but the main reason is to write `packrat/packrat.lock`
-#'
-#' ### Second step: actually build the image
-#'
-#' Now we can use packrat.lock to restore (i.e., install) packages
-#' during the next (real) docker build, using the `Dockerfile`.
-#'
-#' ```
-#' docker build . -t joelnitta/pleurosoriopsis
-#' ```
-#'
-#' ### Third step: rinse, repeat
-#'
-#' Edit the packages below as needed to add new packages
-#' or update old ones (by installing), and repeat Steps 1 and 2.
-#'
-#' To only update new packages instead of running this entire script,
-#' install new packages IN THE CONTAINER using e.g.
-#' install.packages("visNetwork", lib = "/packrat/lib/x86_64-pc-linux-gnu/3.5.3", repos = "https://cran.rstudio.com/")
-#' (The `lib` argument ensures it installs to the packrat library).
-#' The run packrat::snapshot() again.
-#' (You should still add the package to the package list below so that the
-#' whole thing can be built from scratch if needed).
-
-################################################################################
+# This script writes packrat.lock for installing R packages to a docker image.
+# It should be run from within the rocker/verse:3.6.0 container.
+# For more info on installing R packages to docker images with
+# packrat, see https://www.joelnitta.com/post/docker-and-packrat/
 
 ### Initialize packrat ###
 

### 2019-06-20 install_packages.R
Add cowplot, specify commit for gt

@@ -30,6 +30,7 @@ cran_packages <- c(
   "assertthat",
   "broom",
   "conflicted",
+  "cowplot",
   "drake",
   "glue",
   "ggridges",
@@ -48,11 +49,14 @@ install.packages(cran_packages)
 ### Install github packages ###
 github_packages <- c(
   "joelnitta/jntools",
-  "rstudio/gt",
   "thomasp85/patchwork"
 )
 
 remotes::install_github(github_packages)
+
+# gt needs to be installed at a specific commit to deal with a bug
+# https://github.com/rstudio/gt/issues/280
+remotes::install_github("rstudio/gt", ref = "51a812ba6a10769bd24e01c82a3e1b7de44a5a40")
 
 ### Take snapshot ###
 

### 2019-06-20 ms.Rmd
Get rmd to render to pdf
- Use knitr::include_graphics for external images.
- Use specific commit of gt to avoid latex bugs
- Use xelatex engine
- Add page breaks

@@
 ---
-title: "Growth Dynamics of the Independent Gametophytes of *Pleurosoriopsis makinoi* (Polypodiaceae)"
-output: html_document
+title: Growth Dynamics of the Independent Gametophytes of *Pleurosoriopsis makinoi*
+  (Polypodiaceae)
+output:
+  pdf_document:
+    latex_engine: xelatex
 editor_options:
@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
+knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, cache = TRUE)
 
@@
 
+```{r, results='asis', echo = FALSE, eval=(knitr::opts_knit$get('rmarkdown.pandoc.to') == 'latex')}
+# Add pagebreak if rendering to pdf
+cat('\\pagebreak')
+```
+
 ## Figures
 
-```{r fig-1, fig.width=7, fig.height=8, fig.cap = "Fig.
-1.
-The quadrats for measurement of gametophyte cover of *Pleurosoriopsis makinoi* in Unazawa Valley, Okutama-machi, Tokyo, Japan.
+```{r fig-1, fig.width=7, fig.height=7, fig.cap = "The quadrats for measurement of gametophyte cover of *Pleurosoriopsis makinoi* in Unazawa Valley, Okutama-machi, Tokyo, Japan.
 a.
@@
 A 10 cm × 10 cm quadrat frame."}
-
-magick::image_read_pdf(here("ms/Fig1_quadrats.pdf"))
-
-```
-
-```{r fig-2, fig.width=8, fig.height=7, fig.cap = "Fig.
-2.
-Gametophytes of *Pleurosoriopsis makinoi* collected in Unazawa Valley, Okutama-machi, Tokyo, Japan.
+knitr::include_graphics(here("images/Fig1_quadrats.pdf"))
+
+```
+
+```{r fig-2, fig.width=7, fig.height=7, fig.cap = "Gametophytes of *Pleurosoriopsis makinoi* collected in Unazawa Valley, Okutama-machi, Tokyo, Japan.
 a.
@@
 Scale bars: 2 mm for a, 200 μm for b and d, 1 mm for c and e, 100 μm for f."}
-
-magick::image_read_pdf(here("ms/Fig2_gemma_sexualorgans.pdf"))
-
-```
-
-```{r fig-3, fig.width=8, fig.height=7, fig.cap="Fig.
-3.
-Change in growth of *Pleurosoriopsis makinoi* gametophytes observed from May 2009 to May 2014.
+knitr::include_graphics(here("images/Fig2_gemma_sexualorgans.pdf"))
+```
+
+```{r fig-3, fig.width=8, fig.height=7, fig.cap="Change in growth of *Pleurosoriopsis makinoi* gametophytes observed from May 2009 to May 2014.
 a.
@@
 
-```{r fig-4, fig.width=8, fig.height=7, fig.cap = "Fig.
-4.
-Change in microclimate at site with *Pleurosoriopsis makinoi* independent gametophytes (Okutama) and sporophytes (Uratakao) from 2013 to 2014.
+```{r fig-4, fig.width=8, fig.height=7, fig.cap = "Change in microclimate at site with *Pleurosoriopsis makinoi* independent gametophytes (Okutama) and sporophytes (Uratakao) from 2013 to 2014.
 a--b.
@@
 
-```{r fig-5, evaluate = FALSE, fig.width=7, fig.height=8, fig.cap="Fig.
-5.
-Density plots showing difference between microclimate at *Pleurosoriopsis makinoi* independent gametophyte site (Okutama) and sporophyte site (Uratakao).
+```{r fig-5, evaluate = FALSE, fig.width=7, fig.height=8, fig.cap="Density plots showing difference between microclimate at *Pleurosoriopsis makinoi* independent gametophyte site (Okutama) and sporophyte site (Uratakao).
 Significant differences (paired t-test) shown by asterisks: \\*\\*\\*, *p* < 0.001; \\*\\*, *p* < 0.01; \\*, *p* < 0.05.
@@
 
-ggsave(glue::glue("fig-5_{Sys.Date()}.pdf"), width = 7, height = 8, units = "in")
-```
-
-```{r fig-6, fig.width=7, fig.height=8, fig.cap="Fig.
-6.
-Relationships between growth of *Pleurosoriopsis makinoi* gametophytes (total cover, mean gemma number, and mean gemma length) and microclimate at the Okutama site.
+```
+
+```{r fig-6, fig.width=7, fig.height=8, fig.cap="Relationships between growth of *Pleurosoriopsis makinoi* gametophytes (total cover, mean gemma number, and mean gemma length) and microclimate at the Okutama site.
 Microclimatic variables are monthly means of daily light integral (DLI) (mol m^-2^ day^-1^), daily minimum relative humidity (%), and daily mean temperature (°C).
@@
 
-```{r fig-7, fig.width=7, fig.height=4, fig.cap = "Fig.
-7.
-Sizes of gametophytes with antheridia (male), archegonia (female), or no reproductive structures (asexual).
+```{r fig-7, fig.width=7, fig.height=4, fig.cap = "Sizes of gametophytes with antheridia (male), archegonia (female), or no reproductive structures (asexual).
 a.
@@
 
-```{r fig-8, fig.width=7, fig.height=8, fig.cap = "Fig.
-8.
-A comparison of vegetation of the independent gametophyte site with that of a sporophyte site of Pleurosoriopsis.
+```{r fig-8, fig.width=7, fig.height=7, fig.cap = "A comparison of vegetation of the independent gametophyte site with that of a sporophyte site of Pleurosoriopsis.
 a--b.
@@
 "}
-
-magick::image_read_pdf(here("ms/Fig8_vegetation.pdf"))
-
+knitr::include_graphics(here("images/Fig8_vegetation.pdf"))
+```
+
+```{r, results='asis', echo = FALSE, eval=(knitr::opts_knit$get('rmarkdown.pandoc.to') == 'latex')}
+# Add pagebreak if rendering to pdf
+cat('\\pagebreak')
 ```
@@
     var = "",
-    statistic = "t",
-    pval = "p",
+    statistic = md("*t*"),
+    pval = md("*p*"),
     parameter = "df",
@@
   tab_header(
-    title = "Table 1"
-  )
-
+    title = "Table 1",
+    subtitle = md("&nbsp;") # work-around for https://github.com/rstudio/gt/issues/197
+  )
+
+```
+
+```{r, results='asis', echo = FALSE, eval=(knitr::opts_knit$get('rmarkdown.pandoc.to') == 'latex')}
+# Add pagebreak if rendering to pdf
+cat('\\pagebreak')
 ```
@@
     y_var = "",
-    r.squared = "r-squared",
-    adj.r.squared = "adj. r-squared",
-    statistic = "t",
-    p.value = "p",
-    df.residual = "df residual"
+    r.squared = md("r2"),
+    adj.r.squared = md("adj. r2"),
+    statistic = md("*t*"),
+    p.value = md("*p*"),
+    df.residual = "df resid."
   ) %>%
@@
   tab_header(
-    title = "Table 2"
-  )
-
-```
+    title = "Table 2",
+    subtitle = md("&nbsp;") # work-around for https://github.com/rstudio/gt/issues/197
+  )
+
+```

### 2019-06-20 code/packages.R
Delete old commented-out code

@@ -17,7 +17,6 @@ library(patchwork)
 library(readxl)
 library(scales)
 library(tidyverse)
-# library(ggridges)
 
 # Resolve conflicts
 col_factor <- conflict_prefer("col_factor", "readr")

### 2019-06-20 code/plan.R
Use hash without apostrophe for comments

@@ -1,34 +1,34 @@
-#' Workflow Plans
-#'
-#' ## Data sources
-#' 
-#' The raw data are in `.xlsx` files. Some of the data are pre-processed
-#' (mean and sd of count and length of gemmae, 30 min averages of PPFD). 
-#' 
-#' The files include some additional analsyes and other data that won't 
-#' be used here.
+# Workflow Plans
+#
+# ## Data sources
+# 
+# The raw data are in `.xlsx` files. Some of the data are pre-processed
+# (mean and sd of count and length of gemmae, 30 min averages of PPFD). 
+# 
+# The files include some additional analsyes and other data that won't 
+# be used here.
 
 plan <- drake_plan (
   
   # Load and clean data ----
   
-  #' ### Microclimate data
-  #'
-  #' Microclimate variables include PPFD (photon flux density, in μmol of 
-  #' light per sq m per sec), rel. humidity (%), and temperature (°C) 
-  #' measured once every 30 min. 
-  #' 
-  #' PPFD is 30 min-averages of values taken every 4 minutes
-  #' (raw 4 minute values not included).
-  #' 
-  #' There are two sites, Okutama (site of the independent gametophyte
-  #' colony) and Uratakao (site of the sporophyte population). Additional
-  #' sites were also measured, but not included in this analysis
-  #' because of too much missing data due to mechanical failures.
-  #'
-  #' For Okutama, the raw data for temperature, rel. hum,
-  #' and light (PPFD) are in different columns
-  #' in the `xlsx` file, so read in each variable separately.
+  # ### Microclimate data
+  #
+  # Microclimate variables include PPFD (photon flux density, in μmol of 
+  # light per sq m per sec), rel. humidity (%), and temperature (°C) 
+  # measured once every 30 min. 
+  # 
+  # PPFD is 30 min-averages of values taken every 4 minutes
+  # (raw 4 minute values not included).
+  # 
+  # There are two sites, Okutama (site of the independent gametophyte
+  # colony) and Uratakao (site of the sporophyte population). Additional
+  # sites were also measured, but not included in this analysis
+  # because of too much missing data due to mechanical failures.
+  #
+  # For Okutama, the raw data for temperature, rel. hum,
+  # and light (PPFD) are in different columns
+  # in the `xlsx` file, so read in each variable separately.
   
   okutama_temp_raw = read_excel(
[…]

### 2019-06-20 code/plan.R
Render ms to pdf

@@ -225,8 +225,6 @@ plan <- drake_plan (
   # Write out MS
   ms = rmarkdown::render(
     knitr_in("ms.Rmd"),
-    output_format = "html_document",
-    output_file = file_out("ms.html"),
     quiet = TRUE) 
   
 )

### 2019-06-21 make.R
Remove lock_envir = FALSE gt tables now compiling correctly

@@ -25,4 +25,4 @@ source("code/functions.R")
 source("code/plan.R")
 
 # Run analyses ----
-make(plan, lock_envir = FALSE)
+make(plan)

### 2019-06-21 make.R
Remove old comment

@@ -1,4 +1,4 @@
-# make.R 
+# make.R
 
 # Master script for running Pleurorosiopsis analyses.
 # This project uses drake to manage workflows.
@@ -17,11 +17,7 @@ source("code/packages.R")
 pkgconfig::set_config("drake::strings_in_dots" = "literals")
 
 # Load functions and plans  ----
-
-# General functions
 source("code/functions.R")
-
-# Data processing
 source("code/plan.R")
 
 # Run analyses ----

### 2022-10-07 code/plan.R
Remove extra whitespace

@@ -1,26 +1,26 @@
 # Workflow Plans
 #
 # ## Data sources
-# 
+#
 # The raw data are in `.xlsx` files. Some of the data are pre-processed
-# (mean and sd of count and length of gemmae, 30 min averages of PPFD). 
-# 
-# The files include some additional analsyes and other data that won't 
+# (mean and sd of count and length of gemmae, 30 min averages of PPFD).
+#
+# The files include some additional analsyes and other data that won't
 # be used here.
 
 plan <- drake_plan (
-  
+
   # Load and clean data ----
-  
+
   # ### Microclimate data
   #
-  # Microclimate variables include PPFD (photon flux density, in μmol of 
-  # light per sq m per sec), rel. humidity (%), and temperature (°C) 
-  # measured once every 30 min. 
-  # 
+  # Microclimate variables include PPFD (photon flux density, in μmol of
+  # light per sq m per sec), rel. humidity (%), and temperature (°C)
+  # measured once every 30 min.
+  #
   # PPFD is 30 min-averages of values taken every 4 minutes
   # (raw 4 minute values not included).
-  # 
+  #
   # There are two sites, Okutama (site of the independent gametophyte
   # colony) and Uratakao (site of the sporophyte population). Additional
   # sites were also measured, but not included in this analysis
@@ -29,30 +29,30 @@ plan <- drake_plan (
   # For Okutama, the raw data for temperature, rel. hum,
   # and light (PPFD) are in different columns
   # in the `xlsx` file, so read in each variable separately.
-  
+
   okutama_temp_raw = read_excel(
     file_in("data_raw/datalogger_raw.xlsm"),
     sheet = 1,
     range = "A2:B28683",
     col_names = c("date_time", "temp")),
-  
+
   okutama_rh_raw = read_excel(
     file_in("data_raw/datalogger_raw.xlsm"),
     sheet = 1,
     range = "C2:D28683",
     col_names = c("date_time", "rh")),
-  
+
   okutama_par_raw = read_excel(
     file_in("data_raw/datalogger_raw.xlsm"),
     sheet = 1,
     range = "F2:G28204",
     col_names = c("date_time", "par")),
-  
+
   # Combine raw okutama microclimate variables into a single tibble.
   okutama_microclimate_raw =
     full_join(okutama_temp_raw, okutama_rh_raw) %>%
     full_join(okutama_par_raw),
-  
+
[…]

### 2022-10-07 _targets.R
WIP add _targets.R

@@ -1,0 +1,236 @@
+
+library(targets)
+library(tarchetypes)
+source("code/packages.R")
+
+# Workflow Plans
+#
+# ## Data sources
+#
+# The raw data are in `.xlsx` files. Some of the data are pre-processed
+# (mean and sd of count and length of gemmae, 30 min averages of PPFD).
+#
+# The files include some additional analsyes and other data that won't
+# be used here.
+
+
+tar_plan(
+
+  # Load and clean data ----
+
+  # ### Microclimate data
+  #
+  # Microclimate variables include PPFD (photon flux density, in μmol of
+  # light per sq m per sec), rel. humidity (%), and temperature (°C)
+  # measured once every 30 min.
+  #
+  # PPFD is 30 min-averages of values taken every 4 minutes
+  # (raw 4 minute values not included).
+  #
+  # There are two sites, Okutama (site of the independent gametophyte
+  # colony) and Uratakao (site of the sporophyte population). Additional
+  # sites were also measured, but not included in this analysis
+  # because of too much missing data due to mechanical failures.
+  #
+  # For Okutama, the raw data for temperature, rel. hum,
+  # and light (PPFD) are in different columns
+  # in the `xlsx` file, so read in each variable separately.
+
+  okutama_temp_raw = read_excel(
+    file_in("data_raw/datalogger_raw.xlsm"),
+    sheet = 1,
+    range = "A2:B28683",
+    col_names = c("date_time", "temp")),
+
+  okutama_rh_raw = read_excel(
+    file_in("data_raw/datalogger_raw.xlsm"),
+    sheet = 1,
+    range = "C2:D28683",
+    col_names = c("date_time", "rh")),
+
+  okutama_par_raw = read_excel(
+    file_in("data_raw/datalogger_raw.xlsm"),
+    sheet = 1,
+    range = "F2:G28204",
+    col_names = c("date_time", "par")),
+
+  # Combine raw okutama microclimate variables into a single tibble.
+  okutama_microclimate_raw =
+    full_join(okutama_temp_raw, okutama_rh_raw) %>%
+    full_join(okutama_par_raw),
+
+  # Uratakao raw data have a single column for time of the three microclimate
+  # variables, so these can be read in all at once.
+  takao_microclimate_raw = read_excel(
+    file_in("data_raw/datalogger_raw.xlsm"),
+    sheet = 2,
+    range = "D2:G21459",
[…]

### 2022-10-08 _targets.R
Update gitignore
don't ignore .Rprofile

@@ -2,6 +2,7 @@
 library(targets)
 library(tarchetypes)
 source("code/packages.R")
+source("code/functions.R")
 
 # Workflow Plans
 #
@@ -37,19 +38,19 @@ tar_plan(
   # in the `xlsx` file, so read in each variable separately.
 
   okutama_temp_raw = read_excel(
-    file_in("data_raw/datalogger_raw.xlsm"),
+    "data_raw/datalogger_raw.xlsm",
     sheet = 1,
     range = "A2:B28683",
     col_names = c("date_time", "temp")),
 
   okutama_rh_raw = read_excel(
-    file_in("data_raw/datalogger_raw.xlsm"),
+    "data_raw/datalogger_raw.xlsm",
     sheet = 1,
     range = "C2:D28683",
     col_names = c("date_time", "rh")),
 
   okutama_par_raw = read_excel(
-    file_in("data_raw/datalogger_raw.xlsm"),
+    "data_raw/datalogger_raw.xlsm",
     sheet = 1,
     range = "F2:G28204",
     col_names = c("date_time", "par")),
@@ -62,7 +63,7 @@ tar_plan(
   # Uratakao raw data have a single column for time of the three microclimate
   # variables, so these can be read in all at once.
   takao_microclimate_raw = read_excel(
-    file_in("data_raw/datalogger_raw.xlsm"),
+    "data_raw/datalogger_raw.xlsm",
     sheet = 2,
     range = "D2:G21459",
     col_names = c("date_time", "par", "temp", "rh")),
@@ -73,7 +74,7 @@ tar_plan(
   # measured once per month for
   # four 10 x 10 cm plots at the Okutama site.
   cover_raw = read_excel(
-    file_in("data_raw/pleurosoriopsis_data_figs.xlsx"),
+    "data_raw/pleurosoriopsis_data_figs.xlsx",
     sheet = 1,
     skip = 1,
     col_names = c("date", "q_1", "q_2", "q_3", "q_4")
@@ -86,12 +87,12 @@ tar_plan(
   # already cacluated.
 
   gemmae_count_raw = read_xlsx(
-    file_in("data_raw/pleurosoriopsis_data_figs.xlsx"),
+    "data_raw/pleurosoriopsis_data_figs.xlsx",
     sheet = 3,
     range = "A2:C62"),
 
   gemmae_length_raw = read_xlsx(
[…]

### 2022-10-08 code/packages.R
Update package list, resolve conflicts in .Rprofile

@@ -5,7 +5,6 @@ library(assertthat)
 library(broom)
 library(conflicted)
 library(cowplot)
-library(drake)
 library(ggridges)
 library(glue)
 library(gt)
@@ -18,18 +17,4 @@ library(readxl)
 library(scales)
 library(tidyverse)
 
-# Resolve conflicts
-col_factor <- conflict_prefer("col_factor", "readr")
-collapse   <- conflict_prefer("collapse", "dplyr")
-discard    <- conflict_prefer("discard", "purrr")
-expand     <- conflict_prefer("expand", "tidyr")
-extract    <- conflict_prefer("extract", "magrittr")
-filter     <- conflict_prefer("filter", "dplyr")
-gather     <- conflict_prefer("gather", "tidyr")
-ggsave     <- conflict_prefer("ggsave", "ggplot2")
-here       <- conflict_prefer("here", "here")
-intersect  <- conflict_prefer("intersect", "dplyr")
-lag        <- conflict_prefer("lag", "dplyr")
-Position   <- conflict_prefer("Position", "ggplot2")
-setdiff    <- conflict_prefer("setdiff", "dplyr")
-union      <- conflict_prefer("union", "dplyr")
+# any namespace conflicts are resolved in .Rprofile

### 2022-10-08 ms.Rmd
Fix _targets workflow

@@
 # Load objects from drake workflow into memory
-loadd(selected_vars)
-loadd(combined_morph)
-loadd(daily_microclimate)
-loadd(t_test_results)
-loadd(paired_microclimate)
-loadd(nested_paired_microclimate)
-loadd(combined_monthly_morph)
-loadd(linear_model_results)
-loadd(gameto_size)
+tar_load(c(
+  selected_vars,
+  combined_morph,
+  daily_microclimate,
+  t_test_results,
+  paired_microclimate,
+  nested_paired_microclimate,
+  combined_monthly_morph,
+  linear_model_results,
+  gameto_size)
+)
 ```

### 2022-10-08 _targets.R

@@ -227,12 +227,12 @@ tar_plan(
   linear_model_results = make_climate_morph_lm(
     selected_vars = selected_vars,
     combined_monthly_morph = combined_monthly_morph
-  ) #,
+  ),
 
   # Write out MS
-  # tar_knit(
-  #   ms,
-  #   "ms.Rmd"
-  # )
+  tar_render(
+    ms,
+    "ms.Rmd"
+  )
 
 )

### 2022-10-10 install_latex.R
Add Rscript to install latex packages

@@ -1,0 +1,47 @@
+# Install latex packages using tinytex
+
+# This would happen automatically anyways when rendering the Rmd to pdf, 
+# but requires downloading packages and may not work during updates of Tex Live. 
+# Better to install to the docker image once and keep them there.
+
+tinytex::tlmgr_update()
+
+latex_packages <- c(
+  "amsmath",
+  "latex-amsmath-dev",
+  "iftex",
+  "mathspec",
+  "etoolbox",
+  "euenc",
+  "fontspec",
+  "tipa",
+  "xunicode",
+  "geometry",
+  "auxhook",
+  "bigintcalc",
+  "bitset",
+  "etexcmds",
+  "gettitlestring",
+  "hycolor",
+  "hyperref",
+  "intcalc",
+  "kvdefinekeys",
+  "kvsetkeys",
+  "letltxmacro",
+  "ltxcmds",
+  "pdfescape",
+  "refcount",
+  "rerunfilecheck",
+  "stringenc",
+  "uniquecounter",
+  "zapfding",
+  "pdftexcmds",
+  "infwarerr",
+  "kvoptions",
+  "grffile",
+  "titling",
+  "booktabs",
+  "caption"
+)
+
+tinytex::tlmgr_install(latex_packages)

### 2022-10-10 make.sh
Add make.sh

@@ -1,0 +1,7 @@
+#!usr/bin/bash
+set -e
+
+git clone https://github.com/joelnitta/pleurosoriopsis
+cd pleurosoriopsis
+git checkout targets
+Rscript -e 'targets::tar_make()'

### 2022-10-10 install_packages.R
Delete un-used files

@@ -1,66 +1,0 @@
-# This script writes packrat.lock for installing R packages to a docker image.
-# It should be run from within the rocker/verse:3.6.0 container.
-# For more info on installing R packages to docker images with
-# packrat, see https://www.joelnitta.com/post/docker-and-packrat/
-
-### Initialize packrat ###
-
-# Don't let packrat try to find
-# packages to install itself.
-
-install.packages("packrat", repos = "https://cran.rstudio.com/")
-packrat::init(
-  infer.dependencies = FALSE,
-  enter = TRUE,
-  restart = FALSE)
-
-### Setup repositories ###
-
-# Install packages that install packages.
-install.packages("remotes", repos = "https://cran.rstudio.com/")
-
-# Set repos.
-my_repos <- vector()
-my_repos["CRAN"] <- "https://cran.rstudio.com/"
-options(repos = my_repos)
-
-### Install CRAN packages ###
-cran_packages <- c(
-  "assertr",
-  "assertthat",
-  "broom",
-  "conflicted",
-  "cowplot",
-  "drake",
-  "glue",
-  "ggridges",
-  "here",
-  "jntools",
-  "lubridate",
-  "magrittr",
-  "readxl",
-  "scales",
-  "tidyverse",
-  "visNetwork"
-)
-
-install.packages(cran_packages)
-
-### Install github packages ###
-github_packages <- c(
-  "joelnitta/jntools",
-  "thomasp85/patchwork"
-)
-
-remotes::install_github(github_packages)
-
-# gt needs to be installed at a specific commit to deal with a bug
-# https://github.com/rstudio/gt/issues/280
-remotes::install_github("rstudio/gt", ref = "51a812ba6a10769bd24e01c82a3e1b7de44a5a40")
-
-### Take snapshot ###
-
-packrat::snapshot(
-  snapshot.sources = FALSE,
-  ignore.stale = TRUE,
-  infer.dependencies = FALSE)

### 2022-10-10 make.R

@@ -1,24 +1,0 @@
-# make.R
-
-# Master script for running Pleurorosiopsis analyses.
-# This project uses drake to manage workflows.
-# For more information about drake, see
-# https://ropensci.github.io/drake/
-
-# Setup ----
-
-# Set working directory
-setwd(here::here())
-
-# Load packages
-source("code/packages.R")
-
-# Update drake settings
-pkgconfig::set_config("drake::strings_in_dots" = "literals")
-
-# Load functions and plans  ----
-source("code/functions.R")
-source("code/plan.R")
-
-# Run analyses ----
-make(plan)
